---
title: "Octo_monitoring_data"
author: "Ronen Liberman"
date: "2023-12-20"
output:
  html_document: 
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r initiate-environment, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library("knitr")
library("gridExtra")
library("ggpubr")
library("data.table")
library(glmmTMB)
library(bbmle)
library(performance)
library(ggplot2)
library(plotrix)
library(gridExtra)
library(car)
library(emmeans)
library(DHARMa)
library(RRPP)
library(MuMIn)
library(MASS)
library(lattice)
```

```{r customize plot style , warning=FALSE, include=FALSE}
#font_import()

theme_set(theme_minimal(base_size = 12))
theme_update(
  axis.title.x = element_text(margin = margin(12, 0, 0, 0), color = "black"),
  axis.title.y = element_text(margin = margin(0, 12, 0, 0), color = "black"),
  panel.grid.minor = element_blank(),
  panel.border = element_rect(color = "grey45", fill = NA, linewidth = 1.5),
  panel.spacing = unit(.9, "lines"),
  strip.text = element_text(size = rel(1)),
  #plot.title = element_text(size = rel(1), face = "bold", hjust = .5),
  axis.line = element_line(colour = "black"),
        strip.text.x = element_text(size = 10, colour = "black"),
        axis.text=element_text(color="black", size=10))
```

#Data prep. for modelling
Here I filter the data so I can model density of Total population across years 2012/13 - 2023. The data was prepared in: "Creating_binding_all_datasets.R"

```{r}

dat <- read.csv("~/Postdocing_NSU/Octocoral forests/Manuscript/all_density_and_heights_by_stationID.csv")
head(dat)

# Here i test what should be the rangesof the data depth-groups based on number of rows 
# Filter the data for SPP_CODE == "TOCT"
TOCT_data <- dat %>% 
  filter(SPP_Code == "TOCT")

# Convert Depth to numeric (if not already done)
TOCT_data$Depth <- as.numeric(as.character(TOCT_data$Depth))

# Plot the distribution of data by Depth
ggplot(TOCT_data, aes(x = Depth)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Distribution of Data by Depth for SPP_CODE - TOCT", x = "Depth", y = "Number of Rows") 


# Calculate quartiles for the Depth variable
quartiles <- TOCT_data %>%
  summarise(Q1 = quantile(Depth, 0.25),
            Median = quantile(Depth, 0.5),
            Q3 = quantile(Depth, 0.75))

# Create a frequency table of rows per depth
depth_frequency <- TOCT_data %>%
  group_by(Depth_group = cut(Depth, breaks = c(-Inf, quartiles$Q1, quartiles$Median, quartiles$Q3, Inf))) %>%
  summarise(Frequency = n())

# Rename the Depth_group variable for clarity
depth_frequency$Depth_group <- gsub("\\(|\\]", "", as.character(depth_frequency$Depth_group))


### so it seems that if we make 4 groups we need to have it as <21 , 21-28 , 28-45 and above 45 , all in Feet of course.. 
# lets visualize it 

# Define the depth range groups based on quartiles or custom breaks
#dat_3groups <- dat %>% 
 # mutate(Depth_group = case_when(
#    Depth < 20 ~ "Shallow",
 #   Depth >= 20 & Depth < 45 ~ "Mid", 
  #  Depth >= 45 ~ "Deep"
  #))

#dat_3groups %>% filter(SPP_Code == "TOCT") %>%  ggplot(aes(x = Depth_group, fill = Depth_group)) +
 # geom_bar() +
  # labs(title = "Frequency of Observations in Each Depth Group", x = "Depth Group", y = "Frequency") 

# Assign data groups according to the frequncy found above. 
dat_4groups <- dat %>% 
  mutate(Depth_group = case_when(
    Depth < 21 ~ "Shallow",
    Depth >= 21 & Depth < 28 ~ "Mid-shallow", Depth >= 28 & Depth < 45  ~ "Mid-deep",
    Depth >= 45 ~ "Deep"
  ))

#visualize
dat_4groups %>% filter(SPP_Code == "TOCT") %>%  ggplot(aes(x = Depth_group, fill = Depth_group)) +
  geom_bar() +
  labs(title = "Frequency of Observations in Each Depth Group", x = "Depth Group", y = "Frequency") 
# ok , this makes some sense. 


#Generate the dataset
Total.octo <- dat_4groups %>% filter(SPP_Code == "TOCT") 
# check the number of unique ProjectRegion 
Total.octo %>% 
  distinct(ProjectRegion ) 

# make NPSDT the only DT ProjectRegion 
Total.octo <- Total.octo %>%
  mutate(ProjectRegion = ifelse(ProjectRegion == "NPSDT/FKNMS", "NPSDT", ProjectRegion))

Total.octo %>% 
  distinct(Habitat ) 

#Here I explore a bit the oulliars in my data, sime zero's, some low counts in specific sites 
#test for zero density values
#Looking at density 
Total.zero <- Total.octo %>% filter(density == "0")

#check sites to consider removing 

#Total.octo %>% filter(Site_name == "Broward County A")  # dont omit 
#Total.octo %>% filter(Site_name == "Palm Beach 1")  # not found
#Total.octo %>% filter(Site_name == "Davis Rock")  # not in the data
#Total.octo %>% filter(Site_name == "Texas Rock")  # should not omitted I think

# Creating a list of sites that should be omitted from the datasets 
omit_sites <- c("Martin County 1" ,"Martin County 2","Davis Rock","Palm Beach 1")

# omit sites that do not represent octocoral in Florida 
#Convert SiteID and StationID to factors in the Total.octo dataset
Total.octo <- Total.octo %>%
  filter(!(Site_name %in% omit_sites)) %>% 
 # filter(Year!= "2012") %>% 
  mutate(SiteID = as.factor(SiteID),
         StationID = as.factor(StationID))

# look at the frequency of counts 

hist(Total.octo$density, breaks=50, xlab = "density", main = NULL, freq = TRUE) # this doesnt work for some reason
#round density for avoiding non-integer issues with poisson or nbinom. 
Total.octo$counts <- 10*(Total.octo$density)

#round density for avoiding non-integer issues with poisson or nbinom. 
Total.octo$offset <- 10
```


#Modelling All Florida Count data
counts ~ year, spatial scales, enviroment
Here I model using the nbinom2 dist becuase I still have some zeros in the data.

```{r include=FALSE}
nbinom_glmm_full_model = glmmTMB(counts ~ factor(Year) * Subregion +Depth+Habitat+(1|SiteID/StationID),
                          family = nbinom2, 
                          data = Total.octo)
#nbinom_glmm_full_model <- update(nbinom_glmm_full_model, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depthgroup_interaction <- glmmTMB(counts ~ factor(Year) * Depth+ (1|SiteID/StationID),
                      family = nbinom2, 
                      data = Total.octo)
#nbinom_glmm_model_depthgroup_interaction <- #update(nbinom_glmm_model_depthgroup_interaction, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depthgroup <- glmmTMB(counts ~ factor(Year) * Depth_group + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo)
#nbinom_glmm_model_depthgroup <- update(nbinom_glmm_model_depthgroup, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depth <- glmmTMB(counts ~ factor(Year) * Depth + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo)

#nbinom_glmm_model_Subregion <- glmmTMB(density ~ factor(Year) + Subregion+(1|SiteID),
 #                                family = nbinom2, 
#                                 data = Total.octo)
#nbinom_glmm_model_Subregion <- update(nbinom_glmm_model_Subregion, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_Subregion2 <- glmmTMB(counts ~ factor(Year) * Subregion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo)
  #  nbinom_glmm_model_Subregion2 <- update(nbinom_glmm_model_Subregion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))                    
         
nbinom_glmm_model_Subregion3 <- glmmTMB(counts ~ Year * Subregion+Depth + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo)   
#nbinom_glmm_model_Subregion3 <- update(nbinom_glmm_model_Subregion3, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

#nbinom_glmm_model_Subregion4 <- glmmTMB(counts ~ Year * Subregion+(1|SiteID/StationID),
 #                                family = nbinom2, 
  #                               data = Total.octo)
#nbinom_glmm_model_Subregion4 <- update(nbinom_glmm_model_Subregion4, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbionom_glmm_model_ProjectRegion <- glmmTMB(counts ~ factor(Year) + 
ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo)
#nbionom_glmm_model_ProjectRegion <- update(nbionom_glmm_model_ProjectRegion, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_ProjectRegion2 <- glmmTMB(counts ~ Year * ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo)
#nbinom_glmm_model_ProjectRegion2 <- update(nbinom_glmm_model_ProjectRegion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_ProjectRegion3 <- glmmTMB(counts ~ Year * ProjectRegion+Depth+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo)
#nbinom_glmm_model_ProjectRegion3 <- update(nbinom_glmm_model_ProjectRegion3, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbionom_glmm_model_ProjectRegion4 <- glmmTMB(counts ~ factor(Year) + 
ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo)

nbinom_glmm_model_Year <- glmmTMB(counts ~ factor(Year) + Depth+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo)
#nbinom_glmm_model_Year <- update(nbinom_glmm_model_Year, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_habitat <- glmmTMB(counts ~ factor(Year) * Habitat+(1|SiteID),
                                 family = nbinom2, 
                                 data = Total.octo)

#nbinom_glmm_model_habitat2 <- glmmTMB(counts ~ Year * Habitat+Depth+(1|SiteID),
 #                                family = nbinom2, 
  #                               data = Total.octo)

nbinom_glmm_model_habitat2 <- glmmTMB(counts ~ Year + Habitat+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo)
#nbinom_glmm_model_habitat2 <- update(nbinom_glmm_model_habitat2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_null <- glmmTMB(counts ~  + 1,
                           family = nbinom2, 
                            data = Total.octo)
```

##Model selection
```{r Nodel selection, eval=FALSE, message=TRUE, include=FALSE}
AICtab(nbinom_glmm_full_model,
       nbinom_glmm_model_depthgroup_interaction,
       nbinom_glmm_model_depthgroup,nbinom_glmm_model_depth,
    nbinom_glmm_model_Subregion2,nbinom_glmm_model_Subregion3,
    nbionom_glmm_model_ProjectRegion,nbinom_glmm_model_ProjectRegion2,
    nbinom_glmm_model_ProjectRegion3,nbionom_glmm_model_ProjectRegion4,
    nbinom_glmm_model_Year,
    nbinom_glmm_model_habitat,nbinom_glmm_model_habitat2,
    nbinom_glmm_model_null)

#model subregion2 has potential to become the best
summary(nbinom_glmm_model_Subregion2)
```
#2013-2023
Same models but without 2012 data
```{r Remove 2012 data, echo=FALSE}
#Looking at density 
Total.octo_no_2012 <- dat_4groups %>% filter(SPP_Code == "TOCT") %>% 
mutate(ProjectRegion = ifelse(ProjectRegion == "NPSDT/FKNMS", "NPSDT", ProjectRegion)) %>% # Count the number of unique ProjectRegion 
 filter(!(Site_name %in% omit_sites)) %>% 
  filter(Year!= "2012") %>% 
  mutate(SiteID = as.factor(SiteID),
         StationID = as.factor(StationID) 
  )

#Total.octo %>% 
 # distinct(ProjectRegion ) 

#Total.octo %>% 
 # distinct(Habitat )

Total.octo_no_2012$counts <- 10*(Total.octo_no_2012$density)

hist(Total.octo_no_2012$density, breaks=50, xlab = "density", main = NULL, freq = TRUE) # this doesnt work for some reason

```


```{r 2013-2023 models, {r, message=TRUE, warning=FALSE}
nbinom_glmm_full_model_2013 = glmmTMB(counts ~ factor(Year) * Subregion +Depth+Habitat+(1|SiteID/StationID),
                          family = nbinom2, 
                          data = Total.octo_no_2012)
#nbinom_glmm_full_model <- update(nbinom_glmm_full_model, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depthgroup_interaction_2013 <- glmmTMB(counts ~ factor(Year) * Depth+ (1|SiteID/StationID),
                      family = nbinom2, 
                      data = Total.octo_no_2012)
#nbinom_glmm_model_depthgroup_interaction <- #update(nbinom_glmm_model_depthgroup_interaction, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depthgroup_2013 <- glmmTMB(counts ~ factor(Year) * Depth_group + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
#nbinom_glmm_model_depthgroup <- update(nbinom_glmm_model_depthgroup, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depth_2013 <- glmmTMB(counts ~ Year * Depth + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

#nbinom_glmm_model_Subregion <- glmmTMB(density ~ factor(Year) + Subregion+(1|SiteID),
 #                                family = nbinom2, 
#                                 data = Total.octo)
#nbinom_glmm_model_Subregion <- update(nbinom_glmm_model_Subregion, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_Subregion2_2013 <- glmmTMB(counts ~ factor(Year) * Subregion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
  #  nbinom_glmm_model_Subregion2 <- update(nbinom_glmm_model_Subregion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))                    
         
nbinom_glmm_model_Subregion3_2013 <- glmmTMB(counts ~ Year * Subregion+Depth + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)   
#nbinom_glmm_model_Subregion3 <- update(nbinom_glmm_model_Subregion3, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_Subregion4_2013 <- glmmTMB(counts ~ Year * Subregion+(1|SiteID/StationID),
                                family = nbinom2, 
                               data = Total.octo_no_2012)
#nbinom_glmm_model_Subregion4 <- update(nbinom_glmm_model_Subregion4, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbionom_glmm_model_ProjectRegion_2013 <- glmmTMB(counts ~ factor(Year) + 
ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
#nbionom_glmm_model_ProjectRegion <- update(nbionom_glmm_model_ProjectRegion, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_ProjectRegion2_2013 <- glmmTMB(counts ~ Year * ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
#nbinom_glmm_model_ProjectRegion2 <- update(nbinom_glmm_model_ProjectRegion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_ProjectRegion3_2013 <- glmmTMB(counts ~ Year * ProjectRegion+Depth+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
#nbinom_glmm_model_ProjectRegion3 <- update(nbinom_glmm_model_ProjectRegion3, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))


nbinom_glmm_model_Year_2013 <- glmmTMB(counts ~ factor(Year) + Depth+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
#nbinom_glmm_model_Year <- update(nbinom_glmm_model_Year, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_habitat_2013 <- glmmTMB(counts ~ factor(Year) * Habitat+(1|SiteID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

#nbinom_glmm_model_habitat2 <- glmmTMB(counts ~ Year * Habitat+Depth+(1|SiteID),
 #                                family = nbinom2, 
  #                               data = Total.octo)

nbinom_glmm_model_habitat2_2013 <- glmmTMB(counts ~ Year + Habitat+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_habitat3_2013 <- glmmTMB(counts ~ Year + Habitat+Subregion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_null_2013 <- glmmTMB(counts ~  + 1,
                           family = nbinom2, 
                            data = Total.octo)

#summary(nbinom_glmm_model_Subregion2)         

#model subregion2 has potential to become the best

#AICtab(nbionom_glmm_model_ProjectRegion,nbinom_glmm_model_ProjectRegion_2013)
```


```{r Model selection 2013-2023, message=TRUE, warning=TRUE}

AICtab(nbinom_glmm_full_model_2013,
       nbinom_glmm_model_depthgroup_interaction_2013,nbinom_glmm_model_depthgroup_2013,
       nbinom_glmm_model_depth_2013,
    nbinom_glmm_model_Subregion2_2013, nbinom_glmm_model_Subregion3_2013,nbinom_glmm_model_Subregion4_2013,
    nbionom_glmm_model_ProjectRegion_2013,nbinom_glmm_model_ProjectRegion2_2013,
    nbinom_glmm_model_ProjectRegion3_2013,nbinom_glmm_model_Year_2013,
    nbinom_glmm_model_habitat_2013,nbinom_glmm_model_habitat2_2013,nbinom_glmm_model_habitat3_2013,
    nbinom_glmm_model_null_2013)


#summary(nbinom_glmm_model_habitat)
#summary(nbionom_glmm_model_ProjectRegion)


```
nbinom_glmm_model_Subregion2 appears to be the FAVORITE
Should I take model with 2012 or without? Can I reduce some of the outliars? 

#Validation
Plotting residuals vs. fitted shows more or less the same result.

```{r}
# Check for residual pattern within groups and difference between groups     
library(lattice)
xyplot(residuals(nbinom_glmm_model_Subregion2_2013) ~ fitted(nbinom_glmm_full_model_2013),
       panel=function(x, y){
         panel.xyplot(x, y)
         panel.loess(x, y, span = 3)
         panel.lmline(x, y, lty = 2)  # Least squares broken line
       }
)
```
There is some high variation in the residuals - this probably creating the outliars and dispersion roblems inmymodels. lets try to reduce the outliars by removing some marginals locations. 

### omiting sites
```{r}
unique(Total.octo_no_2012$ProjectRegion)
chk <- filter(Total.octo_no_2012, Site_name == "Red Dun Reef")

head(chk)
filter(Total.octo_no_2012, ProjectRegion == "SECREMP") %>%   
ggplot(aes(x = Year, y = density)) +
  geom_point() +
  facet_wrap(~Site_name) +
  ggtitle("Density vs. Year in Region")

# Assuming your data frame is named 'Total.octo_no_2012'
#site_names_NPSDT <- Total.octo_no_2012 %>%
#  filter(ProjectRegion == "NPSDT") %>%
 # distinct(Site_name)

#more sites to omit 
#omit_sites2 <- "Red Dun Reef"
omit_sites2 <- c("Red Dun Reef","Palmata Patch","Loggerhead Patch", "Texas Rock","Broward County A")

# omit sitesfrom overall database
cleaner_total <- Total.octo_no_2012 %>%
filter(!(Site_name %in% omit_sites2)) 
#View(cleaner_total)

```
Run model again with the cleaner data that has just been created and check for outliars again 
```{r echo=FALSE}

nbinom_glmm_full_model_2013_cleaner = glmmTMB(counts ~ factor(Year) * Subregion +Depth+Habitat+(1|SiteID/StationID),
                          family = nbinom2, 
                          data = cleaner_total)
nbinom_glmm_full_model_2013_cleaner <- update(nbinom_glmm_full_model_2013_cleaner, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_full_model_2013_cleaner <- update(nbinom_glmm_full_model_2013, data = cleaner_total)
nbinom_glmm_model_Subregion2_cleaner <- glmmTMB(counts ~ factor(Year) * Subregion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = cleaner_total)


#running best model residuals vs. full model
xyplot(residuals(nbinom_glmm_full_model_2013_cleaner) ~ fitted(nbinom_glmm_full_model_2013_cleaner),
       panel=function(x, y){
         panel.xyplot(x, y)
         panel.loess(x, y, span = 0.75)
         panel.lmline(x, y, lty = 2)  # Least squares broken line
       }
)

```


#Model validation

```{r}

simulationOutput <- simulateResiduals(nbinom_glmm_full_model_2013_cleaner , n = 2000, plot = TRUE,quantreg = T)

#residuals vs predicted plots with lots of data (n > 2000) create quantile regression and a parametric smoother as default.
#2 options to solve - 1. Use quantreg = T in plot(simulationoutput) to see if that solves
#                     2. Group the data by a factor (e.g. by site) and create new residuals using recalculateresiduals(). Then check residuals plot and test residuals using this

#recalculate residuals as per advice from Florian Hartig, Dharma programmer
#recalculates as too many samples for Dharma to calculate, grouping by site

resids <- recalculateResiduals(simulationOutput, group = cleaner_total$Subregion)
plot(resids) # suggest dispersion problem in the model

testResiduals(resids) #residuals uniform, dispersion significant and problematic, no outliers b
testOutliers(resids, type = "bootstrap") #sign , outliars
testDispersion(simulationOutput) #dispersion fine

```

