---
title: "Octo_monitoring_data"
author: "Ronen"
date: "2023-12-20"
output: html_document:
toc: true
theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r initiate-environment, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
library("ggplot2")
library("knitr")
library("gridExtra")
library("ggpubr")
library("data.table")
library(broom)
library(dplyr)  # For data manipulation
library(lubridate)  # For date handling
library(extrafont)
library("scales")
library(car)
```

```{r customize plot style , warning=FALSE, include=FALSE}
#font_import()

theme_set(theme_minimal(base_size = 12))
theme_update(
  axis.title.x = element_text(margin = margin(12, 0, 0, 0), color = "black"),
  axis.title.y = element_text(margin = margin(0, 12, 0, 0), color = "black"),
  panel.grid.minor = element_blank(),
  panel.border = element_rect(color = "grey45", fill = NA, linewidth = 1.5),
  panel.spacing = unit(.9, "lines"),
  strip.text = element_text(size = rel(1)),
  #plot.title = element_text(size = rel(1), face = "bold", hjust = .5),
  axis.line = element_line(colour = "black"),
        strip.text.x = element_text(size = 10, colour = "black"),
        axis.text=element_text(color="black", size=10))
#,
 # plot.title.position = "plot"
#)
```

This is a markdown documnet for analysing CREMP and SE Cremp octocoral monitoring data

## **Data summary**

```{r load data and set it accordingly, paged.print=TRUE}
#starting  from 2013

#All raw data from data of SECREMP/CREMP/DT CREMP can be used here
Height_data <- read.csv("~/Postdocing_NSU/Octocoral forests/Monitoring data/CREMP_CSV_Files/CREMP_OCTO_RawData_2011-2022.csv")
glimpse(Height_data)

#format data 
Height_data$Date <- format(as.Date(Height_data$Date,format = "%m/%d/%Y"))

head(Height_data)
View(Height_data)

# Summary of colony heights (only the three species that interest me) in the different transects whle keeping the metadata 

Sum_height <- Height_data %>%
  #filter(SPP_Code %in% c("PFLE","GVEN","PAME"))  %>%
  filter(sciName!="Eunicea calyculata") %>% 
  group_by(Year,SiteID,StationID, SPP_Code) %>% 
  mutate(mean_height = mean(Height_cm), sd=sd(Height_cm),se = sd / sqrt(n()), biomass=sum(Height_cm),n= n()) %>% 
  select(-Height_cm) %>% 
  distinct() %>% 
  mutate(sciName = dplyr::recode(sciName,
                                 'Pseudopterogorgia americana' = 'Antillogorgia_americana',
                                 "Eunicea flexuosa" = "Eunicea_flexuosa", 
                                 "Gorgonia ventalina" = "Gorgonia_ventalina", "Pseudoplexaura porosa"="Pseudoplexaura porosa", "Pseudopterogorgia bipinnata" = 'Antillogorgia_bipinnata')) %>% #have similar name to the density
  ungroup()

head(Sum_height)
view(Sum_height)
unique(Sum_height$sciName)

#Get the density data
#2013 only
#density_dat <-  read.csv("~/Postdocing_NSU/Octocoral forests/Monitoring data/Me playing with monitoring data/SECREMP_OCTO_Summaries_2013_Density.csv" ) 

#ALL DATA
density_dat <-  read.csv("~/Postdocing_NSU/Octocoral forests/Monitoring data/CREMP_CSV_Files/CREMP_OCTO_Summaries_2011-2022_Density.csv" ) 

glimpse(density_dat)
view(density_dat)

```
#Population level

With this data I can examine the species level or the population level. For the species by using those 3 species. Lets keep all the data I can for the population level. This is truely a proxy for the pop. becuase I dont have all species, right? but this is at least something that can tell us where this is going..

Here I will work with all species I can.

```{r}
pop.dat <- left_join(Sum_height,density_dat, by = c("Year", "Subregion" ,"Habitat" ,  "SiteID" ,   "Site_name" ,"StationID") )
glimpse(pop.dat)
pop.dat <- pop.dat %>% 
  select(-c("Date.y", "Eunicea_calyculata","Gorgonia_ventalina","Antillogorgia_americana","Eunicea_flexuosa","Pseudoplexaura_porosa")) %>% 
 # select(-c(15:21)) %>% 
  rename(oct.density=15)
#view(pop.dat)

#We can view a part of the data - just a check 
LK2013 <- pop.dat %>% 
  filter(Year == "2013") %>% 
  filter(Subregion == "LK")
glimpse(LK2013)
```

## Biomass 
How does the population biomass change? in terms of proxy for all the population and individual species.

### Cleaning data for correctly calculating Biomass
Now in pop dat I have the heights, biomass and density and of each species in each transect. 
But has the data been collected similarly across years? 
```{r}
site_summary <- Height_data %>%
  group_by(Year, Subregion) %>%
  summarise(num_sites = n_distinct(SiteID), num_stations = n_distinct(StationID))

site_wide <- site_summary %>%
  pivot_wider(names_from = Subregion, values_from = c(num_sites,num_stations))
site_wide
view(site_wide)

# It seems that in 2015-2017 another site has been added in LK . this can skew the results. 
#This site was "Red Dun Reef"
#Therefore, I am removing this site from the analysis of biomass and density. 

#It also seems as 2017 had less locations in the MK - possibly becasue of Hurricanes? 
#but this did not effect the data, if any it is a case of more biomass in less sites.. 

#Also, 2011 has less satiton in the UP and thus should be filtered out too
#biomass of each indv. species in each location using x = years , y is mean of biomass / cm2 
```

### Summarize data
```{r}
#rm(biomass_means)
biomass_means <- pop.dat %>% 
  filter(Site_name != "Red Dun Reef") %>% 
  filter(!(Subregion == "UK" & Year == 2011)) %>% 
  group_by(Year,Subregion,sciName) %>%
  summarize(mean_biomass = mean(biomass), sd = sd(biomass),se = sd / sqrt(n()), n=n() ) %>% 
  ungroup()
#view(biomass_means)

total_biomass <- pop.dat %>% 
  filter(Site_name != "Red Dun Reef") %>% 
  filter(!(Subregion == "UK" & Year == 2011)) %>% 
  group_by(Year,Subregion) %>%
  summarize(pooled.taxa_biomass = sum(biomass,na.rm=TRUE), n=n()) %>% 
  ungroup()
view(total_biomass)

#add also pooled data in each subregion. 
# Rename column in total_biomass
total_biomass <- total_biomass %>%
  rename(mean_biomass = pooled.taxa_biomass)
total_biomass$sciName= "pooled taxa" #prepare for binding rows

# Bind rows 
biomass_all <- bind_rows(biomass_means, total_biomass)
# View refined dataframe
view(biomass_all)

# Plot with facet wrap for both sciName and Subregion. Keep plot for pooled taxa seperate because of very different y axis scale. 

#Filter datasets
biomass.pooled <- biomass_all %>% 
  filter(sciName == "pooled taxa")
#%>% mutate(Year= as.character(Year) )
view(biomass.pooled)
biomass.species <- biomass_all %>%
  filter(sciName != "pooled taxa")
```

###Plotting Biomass
```{r}
# Plot for pooled - 
#this is an estimation as I dont have the heights of all the species in each 
#transect and therefore I can only estimate based on the species that have been monitored.

p_biomass.pooled<-  
  ggplot(biomass.pooled, aes(x = Year, y = mean_biomass)) +
  geom_point() +
  geom_path()+
 # geom_errorbar(aes(ymin = mean_biomass - se, 
  #                  ymax = mean_biomass + se)) + #na at thme moment
  scale_x_continuous(breaks=seq(2012,2022,2))+
  facet_wrap(~Subregion) +
  labs(x = "Subregion",
       y = "Biomass proxy",title = "Pooled Taxa Biomass (of 5 species) by Subregion in Florida keys using CREMP data")
p_biomass.pooled

 # Plot for each species
#filter(sciName!='Antillogorgia_bipinnata')  # optional to remove species from this analysis
indv_species_biomass<-  ggplot()+
  geom_jitter(data=pop.dat, aes(x=Year, y = biomass), alpha=0.4)+
  geom_line(data=biomass.species, aes(x = Year, y = mean_biomass, color = sciName ), linewidth = 2  )+
  scale_x_continuous(breaks=seq(2011,2022,2))+
  facet_wrap(~sciName + Subregion, nrow=5) +labs(x = "Year",
                                         y = "Biomass proxy cm2",
                                         title = "Biomass of octocorals by subregion in Florida KEYS")+
  guides(fill = guide_legend()) +
  theme(legend.direction = "horizontal",strip.background = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(color = "black", size = 10))
indv_species_biomass

ggsave(filename = "Biomass_CREMP_pooled species.png", plot = p_biomass.pooled,  width =30 , height = 25,dpi=600,  units = "cm")

ggsave(filename = "Biomass_CREMP_indv_species.png", plot = indv_species_biomass,  width =25 , height =35,dpi=600,  units = "cm")

```

# Density 
How does the population density change? in terms of proxy for all the population and individual species.
I need to create a dataframe with those two df combined. for that we need to make density dat a long df


### Cleaning & summerising the data
```{r}
#I need to create a dataframe with those two df combined. for that we need to make density dat a long df

#step 1: make long 
density_long_dat <-  density_dat %>% pivot_longer(cols=c('Gorgonia_ventalina', 'Antillogorgia_americana', 'Eunicea_flexuosa','Antillogorgia_bipinnata', 'Pseudoplexaura_porosa' ,'Total_Octocorals'),
                                                  names_to='sciName',
                                                  values_to='density')
#step 2: combine data
combined <- left_join(density_long_dat, Sum_height,by = c("Year", "Subregion" ,"Habitat" ,  "SiteID" ,   "Site_name" ,"StationID","sciName" ) )
combined <- combined %>% #filter problematic data points
filter(Site_name != "Red Dun Reef") %>% 
  filter(!(Subregion == "UK" & Year == 2011))
#view(combined)
glimpse(combined)
#Sumemrize data
Sum_density <- combined %>%
  select(-c(Eunicea_calyculata,Date.y)) %>% 
  #filter(SPP_Code %in% c("PFLE","GVEN","PAME"))  %>%
  #  group_by(SiteID,StationID, SPP_Code) %>% 
  group_by(Year,sciName,Subregion) %>%
  summarize(mean_density = mean(density), sd = sd(density),se = sd / sqrt(n()), n=n() )
view(Sum_density)

#Separate the summary by pooled and unpoold (species only). 
dense_pooled <- Sum_density %>% 
  filter(sciName == "Total_Octocorals")

dense_species <-Sum_density %>% 
  filter(sciName != "Total_Octocorals")
#view(dense_pooled)

#seperate the all the dataset with relevant long data and pooled/non pooled
density_CREMP_species <- combined %>% 
  #select(-c(Date.x, Eunicea_calyculata,Antillogorgia_bipinnata,Pseudoplexaura_porosa, Date.y)) %>% 
  filter(sciName != "Total_Octocorals")

density_CREMP_pooled <- combined %>% 
  #select(-c(Date.x, Eunicea_calyculata,Antillogorgia_bipinnata,Pseudoplexaura_porosa, Date.y)) %>% 
  filter(sciName == "Total_Octocorals")
```

###Plotting density
```{r}
total_species_density <-  ggplot()+
 geom_jitter(data=density_CREMP_species, aes(x=Year, y = density), alpha=0.4)+
  geom_line(data=dense_species, aes(x = Year, y = mean_density, color = sciName ), linewidth = 2  )+
  scale_x_continuous(breaks=seq(2011,2022,2))+
  facet_wrap(~sciName + Subregion, nrow=5) +labs(x = "Year",
                                         y = "Colonies m2",
                                         title = "Mean density of octocoral spp. by subregion in Florida KEYS using CREMP data")+
  guides(fill = guide_legend()) +
  theme(legend.direction = "horizontal",strip.background = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(color = "black", size = 10))
total_species_density

#Now for pooled data

total_pool_density <-  ggplot()+
  geom_jitter(data=density_CREMP_pooled, aes(x=Year, y = density), color="orange", alpha=0.4)+
  geom_line(data=dense_pooled, aes(x=Year, y = mean_density), color="skyblue" , linewidth = 2  )+
  scale_x_continuous(breaks=seq(2011,2022,2))+
  facet_wrap(~sciName + Subregion) +labs(x = "Year",
                                         y = "Colonies m2",
                                         title = "Mean density of total octocorals by subregion in Florida KEYS using CREMP data")+
  guides(fill = guide_legend()) +
  theme(legend.direction = "horizontal",strip.background = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(color = "black", size = 10))

total_pool_density

#export *************
ggsave(filename = "Density_CREMP_pooled species.png", plot = total_pool_density,  width =30 , height = 25,dpi=600,  units = "cm")

ggsave(filename = "Density_CREMP_indv_species.png", plot = total_species_density,  width =30 , height = 25,dpi=600,  units = "cm")

```

# Relative abundance - how much of the density does the monitoring species data can explain? 

###Cleaning and summersing the data
```{r}
eco_sum_density <- density_long_dat %>%
  #select(-c(Eunicea_calyculata,Date.y)) %>% 
  #filter(SPP_Code %in% c("PFLE","GVEN","PAME"))  %>%
  #  group_by(SiteID,StationID, SPP_Code) %>% 
  group_by(Year,sciName,Subregion,Site_name) %>%
  summarize(mean_density = mean(density), sd = sd(density),se = sd / sqrt(n()), n=n() )
view(eco_sum_density)

unique(density_long_dat$sciName)

eco_sum_species <- eco_sum_density %>% 
  filter(sciName != "Total_Octocorals") 

eco_sum_pooled <- eco_sum_density %>% 
  filter(sciName == "Total_Octocorals") 
eco_sum_pooled <- eco_sum_pooled%>% 
  rename(mean_all = mean_density, 
         sd_all = sd,  
         se_all = se,
         n_all = n)

glimpse(eco_sum_pooled)

join_density <- left_join(eco_sum_density, eco_sum_pooled,
                          by = c("Year", "Subregion" ,"Site_name") )

join_density <- join_density %>% 
  mutate(rel_abun = mean_density/mean_all) %>% 
  filter(sciName.x!="Total_Octocorals")

# How much of the total abundance does the CREMP data explain?
rel.explanied <- join_density%>%
  #filter(sciName.y!="Total_Octocorals") %>% 
  group_by(Year,Subregion,Site_name,) %>%
  summarize(explined_rel = sum(rel_abun)) 
glimpse(rel.explanied)  

#rel.explanied.med <- rel.explanied %>%
 # group_by(Year, Subregion) %>%
  #summarise(med = quantile(explined_rel, 0.5,na.rm= T))
```

###Plotting relative abundace
```{r}
p_pooled_rel <- ggplot(rel.explanied, aes(x = Year, y = explined_rel, group= Year)) +
  geom_boxplot() +
  #  geom_text(aes(label = round(med, 2))) +
  facet_wrap(~Subregion) +
  #ggplot(rel.explanied, aes(x = Year, y = explined_rel)) +
  # Violin with median line
  #geom_violin(alpha = 0.2, draw_quantiles = c(0.5)) +
  scale_x_continuous(breaks=seq(2011,2022,2))+
  # Points for each site
  #geom_jitter(aes(fill = Site_name), 
  #position = position_jitter(width = 0.2),
  #alpha = 0.5) +
  # Remove Subregion color mapping
  #guides(color = "none") + 
  facet_wrap(~Subregion) +
  labs(
    title = "Explained Relative Abundance by Subregion in the KEYS using CREMP data",
    x = "Year",
    y = "Explined Relative Abundance"
  )
p_pooled_rel
#~ 70% of the total abundance is explained by 5 species - not bad, not perfect , and ton of room for improvment 
#Variability in UK is higher. 
#Relative abundances by species: Look at each species as a proportion of total abundance per year. 
#Analyze how the relative abundances change over time.
# Analyze relative abundance per species
join_density_sum <- join_density%>%
  #filter(sciName.y!="Total_Octocorals") %>% 
  group_by(Year,sciName.x,Subregion) %>%
  summarize(mean_rel = mean(rel_abun, na.rm=T), sd_rel = sd(rel_abun, na.rm=T)) 

p_rel_density <- ggplot(data=join_density_sum, aes(x = Year, y = mean_rel)) + 
  geom_jitter(data = join_density, aes(x = Year, y = rel_abun, color = sciName.x), alpha = 0.5)+
  geom_path(linewidth=1) +
  facet_wrap(~sciName.x+Subregion, ncol=3)+
  scale_x_continuous(breaks=seq(2011,2022,2))+ 
  labs(x = "Year",
       y = "Mean relative abundance",
       title = "Mean relative abundance of octocorals by subregion in Florida KEYS using CREMP data")+
  guides(fill = guide_legend()) +
  theme(legend.direction = "horizontal",strip.background = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(color = "black", size = 10))
p_rel_density

#export *************
ggsave(filename = "Relative Abundance of pooled_taxa_CREMP.jpg", plot = p_pooled_rel,  width =30 , height = 25,dpi=600,  units = "cm")

ggsave(filename = "Relative Abundance of species_CREMP.jpg", plot = p_rel_density,width =25 , height = 35,dpi=600,  units = "cm")

```


#Per Site

I want to sort of look at the sites from a trend perspective, has the octocoral density changed significanly in the different sites? which sites show changes and can we relate this to different parametes such as enviroment, depth, anthropogenic stress? 

```{r}
#Calculating linear trends across time per site
###### Pooled and species density trends ########
  
#working on the df "combined" that was created under the density section
  glimpse(combined)
  species_data <- combined %>%
    select(Year, Site_name, sciName, density) %>%
    filter(!is.na(density), !is.na(Year), !is.na(sciName), !is.na(Site_name)) 
 glimpse(species_data)
 str(species_data) 
 #To create a linear model for each combination of species and site, you can use the following R code:
  models <- species_data %>%
    #filter(sciName=="Total_Octocorals") %>% 
    group_by(Site_name) %>%
    do(model = lm(density ~ sciName, data = .)) %>%
    ungroup() %>% 
    transmute(Site_name, SiteCoef = map(model,tidy)) %>%
    unnest(Site_name)
    
  models
  #####################
  
  models <- species_data %>%
    filter(sciName!="Total_Octocorals") %>% 
    group_by(Site_name) %>%
    do(mod = glance(lm(density ~ Year, data = .)))%>%
    tidyr::unnest(mod)
  view(models)

#  This is a model for a change in each speceis/ pooled for each year and each site
    ###############################
  
  adjusted_models <- species_data %>%
    filter(sciName=="Total_Octocorals") %>% 
    group_by(Site_name) %>%
    do(mod = tidy(lm(density ~ Year, data = .))) %>%
    unnest(mod)
  
# This is a model only for the pooled 
  
 #In this model I waant to know the rate of in the total density change in each site from the earliest data point to the most recent 
   recent_year_data <- species_data %>%
     filter(sciName=="Total_Octocorals") %>% 
    filter(!is.na(density) & !is.na(Year)) %>%  # Filter out rows with NA values for 'density' and 'Year'
    group_by(Site_name) %>%
    filter(Year %in% c(min(Year), max(Year)))  # Filter for the rows with the earliest and most recent year for each site
  
  # Calculate the total change in species density using the slope coefficient
  model_early_to_recent <- recent_year_data %>%
  group_by(Site_name) %>%
    do(mod = tidy(lm(density ~ Year, data = .))) %>%
     unnest(mod)
    
  # View the model-level summary statistics
  model_early_to_recent
# Summary
   # Assuming your model results are stored in the dataframe 'model_early_to_recent'
    
  # Filter for term = Year and arrange by slope
    year_term_summary <- model_early_to_recent %>%
      filter(term == "Year") %>%
      select(Site_name, term, estimate, p.value) %>%
      arrange(estimate)
   year_term_summary
   
    year_term_summary <-  year_term_summary %>%
      mutate(depth = case_when(
        grepl("Deep", Site_name) ~ "deep",
        grepl("Shallow", Site_name) ~ "shallow",
        TRUE ~ NA_character_
      ))
    
    # Binding metadata about the sites from 'combined' to 'year_term_summary' by 'Site_name'
    year_term_summary <- year_term_summary %>%
      left_join(combined %>% distinct(Site_name, Habitat, Subregion), by = "Site_name")
    
    #So in total, how many sites increase/ decrease or stayed the same? 
    # Generate a new table showing the count of significant increase, decrease, and no change
    # Adding average and SE to each of the 3 categories in the site table
    summary_with_stats <- year_term_summary %>%
      mutate(change_category = case_when(
        estimate > 0 & p.value < 0.05 ~ "Significant Increase",
        estimate < 0 & p.value < 0.05 ~ "Significant Decrease",
        TRUE ~ "No Significant Change"
      )) %>%
      group_by(change_category) %>%
      summarise(Average = mean(estimate), SE = sd(estimate) / sqrt(n())) %>%
      ungroup() %>%
      right_join(site_summary, by = c("change_category" = "Change_Category"))
  
    # View the summary of term = Year and the new table showing the count of change categories
    summary_with_stats # 8/20 have increased! 40% in the Keys. 
```

```{r}
#Plot density by site Name

total_pool_density_in_site <-  ggplot()+
  geom_jitter(data=density_CREMP_pooled, aes(x=Year, y = density), color="orange")+
  #geom_line(data=dense_pooled, aes(x=Year, y = mean_density), color="skyblue" , linewidth = 2  )+
  scale_x_continuous(breaks=seq(2011,2022,2))+
  facet_wrap(~Site_name) +labs(x = "Year",
                                         y = "Colonies m2",
                                         title = "Mean density of total octocorals by Site in Florida KEYS using CREMP data")+
  guides(fill = guide_legend()) +
  theme(legend.direction = "horizontal",strip.background = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(color = "black", size = 10))
total_pool_density_in_site

#Site name by species
head(density_CREMP_species)

library(RColorBrewer)

colors <- brewer.pal(5, "Set2")

scale_color_discrete <- scale_color_manual(values = colors)

total_species_density_By_site <-  ggplot()+
  geom_jitter(data=density_CREMP_species, aes(x=Year, y = density, color = sciName ))+
  #geom_line(data=dense_species, aes(x = Year, y = mean_density, color = sciName ), linewidth = 2  )+
  scale_x_continuous(breaks=seq(2011,2022,2))+
  scale_color_manual(values = colors)+
  facet_wrap(~Subregion+Site_name) +labs(x = "Year",
                                                 y = "Colonies m2",
                        title = "Mean density of octocoral spp. by Site in 
                               Florida KEYS using CREMP data")+
  guides(fill = guide_legend()) +
  theme(legend.direction = "horizontal",strip.background = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(color = "black", size = 10))
total_species_density_By_site

ggsave(filename = "exploring_Density_CREMP_indv_by_Site.jpg", plot = total_species_density_By_site,  width =30 , height = 35,dpi=600,  units = "cm")

```


```{r}

#Can I check trends per site? How many sites I actually have? 
combined %>% 
  group_by(Subregion) %>% 
  tally()
# 4 regions 
combined %>% 
  group_by(Habitat) %>% 
  tally()

# 4 Habitats 

combined %>% 
  group_by(Subregion, Habitat, Site_name) %>% 
  tally()
# not all regions and have all habitats. 

#Thus,
#I think I should start from looking at the sub region level, and represent the habitat too.

# How can I show if there relationship between density and colony height? 

#Maybe a good start will be a see how density of each species and the total changes over sampling times? 

#summerize density 



#plot - Thus would be later compared per year, so year should be my x axis 
mycolo <- c('#d7191c','#252525', '#006d2c','#08519c' )

P_density <- ggplot(combined, aes(x=Year, y= density, color=sciName  ))+
geom_point()+
  scale_color_manual(values = mycolo)+
  facet_grid(.~Subregion)

P_density
```
