---
title: "Modeling Florida octocoral trends: Density data"
author: "Ronen Liberman"
date: "6-17-2024"
output: 
  html_document:
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r initiate-environment, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library("knitr")
library("gridExtra")
library("ggpubr")
library("data.table")
library(glmmTMB)
library(bbmle)
library(performance)
library(ggplot2)
library(plotrix)
library(gridExtra)
library(car)
library(emmeans)
library(DHARMa)
library(RRPP)
library(MuMIn)
library(MASS)
library(lattice)
```

```{r customize plot style , warning=FALSE, include=FALSE}
#font_import()

theme_set(theme_minimal(base_size = 12))
theme_update(
  axis.title.x = element_text(margin = margin(12, 0, 0, 0), color = "black"),
  axis.title.y = element_text(margin = margin(0, 12, 0, 0), color = "black"),
  panel.grid.minor = element_blank(),
  panel.border = element_rect(color = "grey45", fill = NA, linewidth = 1.5),
  panel.spacing = unit(.9, "lines"),
  strip.text = element_text(size = rel(1)),
  #plot.title = element_text(size = rel(1), face = "bold", hjust = .5),
  axis.line = element_line(colour = "black"),
        strip.text.x = element_text(size = 10, colour = "black"),
        axis.text=element_text(color="black", size=10))
```

*In this markdown file I am going to work with data that was prepared in a previous script. 
I am going to clean the data a little bit, then using model selection test for the most significant parameters that influence state of octocoral population in the last 11 years*

#Data prep. for modelling
Here I filter the data so I can model density of Total population across years 2013 - 2023. The data was prepared in: "Creating_binding_all_datasets.R"

```{r}
dat <- read.csv("~/Postdocing_NSU/Octocoral forests/Manuscript/all_density_and_heights_by_stationID.csv")
head(dat)

# Here i test what should be the rangesof the data depth-groups based on number of rows 
# Filter the data for SPP_CODE == "TOCT"
TOCT_data <- dat %>% 
  filter(SPP_Code == "TOCT")

# Convert Depth to numeric (if not already done)
TOCT_data$Depth <- as.numeric(as.character(TOCT_data$Depth))

# Plot the distribution of data by Depth
ggplot(TOCT_data, aes(x = Depth)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Distribution of Data by Depth for SPP_CODE - TOCT", x = "Depth", y = "Number of Rows") 


# Calculate quartiles for the Depth variable
quartiles <- TOCT_data %>%
  summarise(Q1 = quantile(Depth, 0.25),
            Median = quantile(Depth, 0.5),
            Q3 = quantile(Depth, 0.75))

# Create a frequency table of rows per depth
depth_frequency <- TOCT_data %>%
  group_by(Depth_group = cut(Depth, breaks = c(-Inf, quartiles$Q1, quartiles$Median, quartiles$Q3, Inf))) %>%
  summarise(Frequency = n())

# Rename the Depth_group variable for clarity
depth_frequency$Depth_group <- gsub("\\(|\\]", "", as.character(depth_frequency$Depth_group))

# Assign data groups according to the frequncy found above. 
dat_4groups <- dat %>% 
  mutate(Depth_group = case_when(
    Depth < 21 ~ "Shallow",
    Depth >= 21 & Depth < 28 ~ "Mid-shallow", Depth >= 28 & Depth < 45  ~ "Mid-deep",
    Depth >= 45 ~ "Deep"
  ))

#visualize
dat_4groups %>% filter(SPP_Code == "TOCT") %>%  ggplot(aes(x = Depth_group, fill = Depth_group)) +
  geom_bar() +
  labs(title = "Frequency of Observations in Each Depth Group", x = "Depth Group", y = "Frequency") 
# ok , this makes some sense. 

#Generate the dataset
Total.octo <- dat_4groups %>% filter(SPP_Code == "TOCT") 
# check the number of unique ProjectRegion 
Total.octo %>% 
  distinct(ProjectRegion ) 

# make NPSDT the only DT ProjectRegion 
Total.octo <- Total.octo %>%
  mutate(ProjectRegion = ifelse(ProjectRegion == "NPSDT/FKNMS", "NPSDT", ProjectRegion))

Total.octo %>% 
  distinct(Habitat ) 

#Here I explore a bit the outlliars in my data, some zero's, some low counts in specific sites 
#test for zero density values
#Looking at density 
Total.zero <- Total.octo %>% filter(density == "0")

# Creating a list of sites that should be omitted from the datasets 
omit_sites <- c("Martin County 1" ,"Martin County 2","Davis Rock","Palm Beach 1","Red Dun Reef","Palmata Patch","Loggerhead Patch", "Texas Rock","Broward County A")

#omit_sites2 <- c("Red Dun Reef","Palmata Patch","Loggerhead Patch", "Texas Rock","Broward County A")

Total.octo_no_2012 <- dat_4groups %>% filter(SPP_Code == "TOCT") %>% 
mutate(ProjectRegion = ifelse(ProjectRegion == "NPSDT/FKNMS", "NPSDT", ProjectRegion)) %>% # Count the number of unique ProjectRegion
 filter(!(Site_name %in% omit_sites)) %>% 
  filter(Year!= "2012") %>% 
  mutate(SiteID = as.factor(SiteID),
         StationID = as.factor(StationID),
         Year =  as.factor(Year),
         counts = (10*density)
  )

#view(Total.octo_no_2012)
#Total.octo %>% 
 # distinct(ProjectRegion ) 

#Total.octo %>% 
 # distinct(Habitat )

# look at the frequency of counts 

hist(Total.octo_no_2012$counts, breaks=50, xlab = "density", main = NULL, freq = TRUE) 


#round density for avoiding non-integer issues with poisson or nbinom. 
Total.octo_no_2012$offset <- 10
```

Using different models I can examine the most significant factors affecting Total octocoral abundance. I am using count data and different explenatory varables. I am using a model selection method using the AIC parameter to select for the best model. 
Afterwards I am running a model validation functions on the best model. 

Modeling using NBINOM2 dist. 

```{r }
nbinom_glmm_full_model_2013 = glmmTMB(counts ~ Year * Subregion +Depth+Habitat+(1|SiteID/StationID),
                          family = nbinom2, 
                          data = Total.octo_no_2012)

#nbinom_glmm_full_model <- update(nbinom_glmm_full_model, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depthgroup_interaction_2013 <- glmmTMB(counts ~ Year * Depth+ (1|SiteID/StationID),
                      family = nbinom2, 
                      data = Total.octo_no_2012)
#nbinom_glmm_model_depthgroup_interaction <- #update(nbinom_glmm_model_depthgroup_interaction, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depthgroup_2013 <- glmmTMB(counts ~ Year * Depth_group + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
#nbinom_glmm_model_depthgroup <- update(nbinom_glmm_model_depthgroup, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depth_2013 <- glmmTMB(counts ~ Year + Depth + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_Subregion_2013 <- glmmTMB(density ~ Year * Subregion+(1|SiteID/StationID),
                               family = nbinom2, 
                                data = Total.octo_no_2012)

#nbinom_glmm_model_Subregion <- update(nbinom_glmm_model_Subregion, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_Subregion2_2013 <- glmmTMB(counts ~ Year + Subregion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
  #  nbinom_glmm_model_Subregion2 <- update(nbinom_glmm_model_Subregion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))                    
         
nbinom_glmm_model_Subregion3_2013 <- glmmTMB(counts ~ Year * Subregion+Depth + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)   
#nbinom_glmm_model_Subregion3 <- update(nbinom_glmm_model_Subregion3, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))


nbionom_glmm_model_ProjectRegion_2013 <- glmmTMB(counts ~ Year * ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
#nbionom_glmm_model_ProjectRegion <- update(nbionom_glmm_model_ProjectRegion, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_ProjectRegion2_2013 <- glmmTMB(counts ~ Year + ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
#nbinom_glmm_model_ProjectRegion2 <- update(nbinom_glmm_model_ProjectRegion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_ProjectRegion3_2013 <- glmmTMB(counts ~ Year * ProjectRegion+Depth+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_ProjectRegion4_2013 <- glmmTMB(counts ~ Year * ProjectRegion+Habitat+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_Year_2013 <- glmmTMB(counts ~ as.numeric(Year) + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
#nbinom_glmm_model_Year <- update(nbinom_glmm_model_Year, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))


nbinom_glmm_model_habitat_2013 <- glmmTMB(counts ~ Year * Habitat+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)


nbinom_glmm_model_habitat2_2013 <- glmmTMB(counts ~ Year + Habitat+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_habitat3_2013 <- glmmTMB(counts ~ Year + Habitat+Subregion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_null_2013 <- glmmTMB(counts ~  + 1,
                           family = nbinom2, 
                            data = Total.octo_no_2012)
```

```{r }
AICtab(nbinom_glmm_full_model_2013,
       nbinom_glmm_model_depthgroup_interaction_2013,nbinom_glmm_model_depthgroup_2013,
       nbinom_glmm_model_depth_2013,
    nbinom_glmm_model_Subregion_2013,nbinom_glmm_model_Subregion2_2013, nbinom_glmm_model_Subregion3_2013,
    nbionom_glmm_model_ProjectRegion_2013,nbinom_glmm_model_ProjectRegion2_2013,
    nbinom_glmm_model_ProjectRegion3_2013,nbinom_glmm_model_ProjectRegion4_2013,
    nbinom_glmm_model_Year_2013,
    nbinom_glmm_model_habitat_2013,nbinom_glmm_model_habitat2_2013,nbinom_glmm_model_habitat3_2013,
    nbinom_glmm_model_null_2013)
```

**Summary of best model**
The model that explains the variation of count data ~ Year * Habitat is the FAVORITE. 

```{r echo=FALSE, warning=TRUE}
Best_model = nbinom_glmm_model_Subregion_2013
summary(Best_model)
```

Model validation

**Visualizing residuals**
Plotting residuals vs. fitted shows more or less the same result.

```{r}
# Check for residual pattern within groups and difference between groups     
xyplot(residuals(Best_model) ~ fitted(nbinom_glmm_full_model_2013),
       panel=function(x, y){
         panel.xyplot(x, y)
         panel.loess(x, y, span = 3)
         panel.lmline(x, y, lty = 2)  # Least squares broken line
       }
)
```
There is some high variation in the residuals - this probably creating the outliars and dispersion roblems inmymodels. lets try to reduce the outliars by removing some marginals locations. 

Validation of Best model
Here I use the DHarMA package to validate my best model 
QQ plot is on the red line, but still there is a significant result in the uniformity test. 
As it stands, the model is almost validated, but it seems some there is an indication of issues with the uniformity of the data. 

```{r echo=TRUE}
simulationOutput <- simulateResiduals(Best_model, n = 2000, plot = TRUE,quantreg = T)
testDispersion(simulationOutput) #dispersion fine
```

Validation per group - I examine individual groups in the dataset. #recalculating residuals , as I might have too many samples for Dharma to calculate, grouping by site, Year, or Subregion

#Site - this parameter has enough data point and therefore can be a good one to look into. All parameters are non significant

```{r}
resids.site <- recalculateResiduals(simulationOutput, group = Total.octo_no_2012$SiteID)
plot(resids.site) # suggest dispersion problem in the model

testResiduals(resids.site) #residuals uniform, dispersion significant and problematic, no outliers b
testOutliers(resids.site, type = "bootstrap") #sign , outliars
testDispersion(simulationOutput) #dispersion fine

```
##Year - 
This factor shows significant in uniformity, however this is probably becaus ethere is only 11 years in the dataset, however, dispersion and outliars are  OK. 
```{r}
resids.year<- recalculateResiduals(simulationOutput, group = Total.octo_no_2012$Year)
plot(resids.year) # suggest dispersion problem in the model

testResiduals(resids.year) #residuals uniform, dispersion significant and problematic, no outliers b
testOutliers(resids.year, type = "bootstrap") #we dont have outliars

```

Validation per group - Subregion 
Rest of the subgrouos were already shown above

###Subregion - this parameter is significant for dispersion, this is probably due to lack of subregions in the dataset, only 7, however, we dont find outliars and the data is pretty uniform. 
```{r}
resids.sub <- recalculateResiduals(simulationOutput, group = Total.octo_no_2012$Subregion)
plot(resids.sub) # suggest dispersion problem in the model

testResiduals(resids.sub) #residuals uniform, dispersion significant and problematic, no outliers b
testOutliers(resids.sub, type = "bootstrap") #sign , outliars
```


#Second best model - **Habitat **
```{r}
# Check for residual pattern within groups and difference between groups     
xyplot(residuals(nbinom_glmm_model_habitat_2013) ~ fitted(nbinom_glmm_full_model_2013),
       panel=function(x, y){
         panel.xyplot(x, y)
         panel.loess(x, y, span = 3)
         panel.lmline(x, y, lty = 2)  # Least squares broken line
       }
)
```
There is some high variation in the residuals - this probably creating the outliars and dispersion roblems inmymodels. lets try to reduce the outliars by removing some marginals locations. 

Validation of second Best model - Year * Habitat

```{r echo=TRUE}
simulationOutput <- simulateResiduals(nbinom_glmm_model_habitat_2013, n = 2000, plot = TRUE,quantreg = T)
testDispersion(simulationOutput) #dispersion fine
```

###Habitat - this parameter is significant for outliars, this is probably due to lack of habitats in the dataset.

```{r}
resids.hab <- recalculateResiduals(simulationOutput, group = Total.octo_no_2012$Habitat)
plot(resids.hab) # suggest dispersion problem in the model

testResiduals(resids.hab) #residuals uniform, dispersion significant and problematic, no outliers b
testOutliers(resids.hab, type = "bootstrap") #sign , outliars
```

#Old stuff from here # dont knit 

```{r Remove 2012 data, eval=FALSE, include=FALSE}
#Looking at density 
Total.octo_no_2012 <- dat_4groups %>% filter(SPP_Code == "TOCT") %>% 
mutate(ProjectRegion = ifelse(ProjectRegion == "NPSDT/FKNMS", "NPSDT", ProjectRegion)) %>% # Count the number of unique ProjectRegion 
 filter(!(Site_name %in% omit_sites)) %>% 
  filter(Year!= "2012") %>% 
  mutate(SiteID = as.factor(SiteID),
         StationID = as.factor(StationID) 
  )

#Total.octo %>% 
 # distinct(ProjectRegion ) 

#Total.octo %>% 
 # distinct(Habitat )

Total.octo_no_2012$counts <- 10*(Total.octo_no_2012$density)

hist(Total.octo_no_2012$density, breaks=50, xlab = "density", main = NULL, freq = TRUE) # this doesnt work for some reason

```
NBINOM1

The difference between using family = nbinom1 and family = nbinom2 is in the parameterization of the negative binomial distribution. ​ In nbinom1, the variance of the distribution is equal to µ(1 + α), where µ is the mean and α is the estimated dispersion parameter. ​ This parameterization assumes that the variance increases with the mean. ​ In nbinom2, the variance of the distribution is µ(1 + µ/θ), where θ is the estimated dispersion parameter. ​ This parameterization assumes that the variance decreases with the mean. ​ The choice between the two parameterizations depends on the specific characteristics of the data and the assumptions that best fit the relationship between the mean and variance.
Nbinom1 - running models with nbinom2 dist. 

```{r eval=FALSE, include=FALSE}
nbinom_glmm_full_model_2013 = glmmTMB(counts ~ Year * Subregion +Depth+Habitat+(1|SiteID/StationID),
                          family = nbinom1, 
                          data = Total.octo_no_2012)

#nbinom_glmm_full_model <- update(nbinom_glmm_full_model, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depthgroup_interaction_2013 <- glmmTMB(counts ~ Year * Depth+ (1|SiteID/StationID),
                      family = nbinom1, 
                      data = Total.octo_no_2012)
#nbinom_glmm_model_depthgroup_interaction <- #update(nbinom_glmm_model_depthgroup_interaction, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depthgroup_2013 <- glmmTMB(counts ~ Year * Depth_group + (1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)
#nbinom_glmm_model_depthgroup <- update(nbinom_glmm_model_depthgroup, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depth_2013 <- glmmTMB(counts ~ Year + Depth + (1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_Subregion_2013 <- glmmTMB(density ~ Year * Subregion+(1|SiteID/StationID),
                               family = nbinom1, offset = offset,
                                data = Total.octo_no_2012)

#nbinom_glmm_model_Subregion <- update(nbinom_glmm_model_Subregion, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_Subregion2_2013 <- glmmTMB(counts ~ Year + Subregion+(1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)
  #  nbinom_glmm_model_Subregion2 <- update(nbinom_glmm_model_Subregion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))                    
         
nbinom_glmm_model_Subregion3_2013 <- glmmTMB(counts ~ Year * Subregion+Depth + (1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)   
#nbinom_glmm_model_Subregion3 <- update(nbinom_glmm_model_Subregion3, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))


nbionom_glmm_model_ProjectRegion_2013 <- glmmTMB(counts ~ Year * ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom1, offset = offset,
                                 data = Total.octo_no_2012)
#nbionom_glmm_model_ProjectRegion <- update(nbionom_glmm_model_ProjectRegion, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_ProjectRegion2_2013 <- glmmTMB(counts ~ Year + ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)
#nbinom_glmm_model_ProjectRegion2 <- update(nbinom_glmm_model_ProjectRegion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_ProjectRegion3_2013 <- glmmTMB(counts ~ Year * ProjectRegion+Depth+(1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_ProjectRegion4_2013 <- glmmTMB(counts ~ Year * ProjectRegion+Habitat+(1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_Year_2013 <- glmmTMB(counts ~ as.numeric(Year) + (1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)
#nbinom_glmm_model_Year <- update(nbinom_glmm_model_Year, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))


nbinom_glmm_model_habitat_2013 <- glmmTMB(counts ~ Year * Habitat+(1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)


nbinom_glmm_model_habitat2_2013 <- glmmTMB(counts ~ factor(Year) + Habitat+(1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_habitat3_2013 <- glmmTMB(counts ~ Year + Habitat+Subregion+(1|SiteID/StationID),
                                 family = nbinom1, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_null_2013 <- glmmTMB(counts ~  + 1,
                           family = nbinom1, 
                            data = Total.octo)

#summary(nbinom_glmm_model_Subregion2)         

#model subregion2 has potential to become the best

#AICtab(nbionom_glmm_model_ProjectRegion,nbinom_glmm_model_ProjectRegion_2013)
```


```{r eval=FALSE, message=TRUE, warning=TRUE, include=FALSE}
AICtab(nbinom_glmm_full_model_2013,
       nbinom_glmm_model_depthgroup_interaction_2013,nbinom_glmm_model_depthgroup_2013,
       nbinom_glmm_model_depth_2013,
    nbinom_glmm_model_Subregion_2013,nbinom_glmm_model_Subregion2_2013, nbinom_glmm_model_Subregion3_2013,
    nbionom_glmm_model_ProjectRegion_2013,nbinom_glmm_model_ProjectRegion2_2013,
    nbinom_glmm_model_ProjectRegion3_2013,nbinom_glmm_model_ProjectRegion4_2013,
    nbinom_glmm_model_Year_2013,
    nbinom_glmm_model_habitat_2013,nbinom_glmm_model_habitat2_2013,nbinom_glmm_model_habitat3_2013,
    nbinom_glmm_model_null_2013)
```



#**List of models:** 
nbinom_glmm_full_model_2013 = glmmTMB(counts ~ Year * Subregion +Depth+Habitat+(1|SiteID/StationID),
                          family = nbinom2, 
                          data = Total.octo_no_2012)


nbinom_glmm_model_depthgroup_interaction_2013 <- glmmTMB(counts ~ Year * Depth+ (1|SiteID/StationID),
                      family = nbinom2, 
                      data = Total.octo_no_2012)


nbinom_glmm_model_depthgroup_2013 <- glmmTMB(counts ~ Year * Depth_group + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_depth_2013 <- glmmTMB(counts ~ Year + Depth + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_Subregion_2013 <- glmmTMB(density ~ Year * Subregion+(1|SiteID/StationID),
                               family = nbinom2, 
                                data = Total.octo_no_2012)


nbinom_glmm_model_Subregion2_2013 <- glmmTMB(counts ~ Year + Subregion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)
  #  nbinom_glmm_model_Subregion2 <- update(nbinom_glmm_model_Subregion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))                    
         
nbinom_glmm_model_Subregion3_2013 <- glmmTMB(counts ~ Year * Subregion+Depth + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)   


nbionom_glmm_model_ProjectRegion_2013 <- glmmTMB(counts ~ Year * ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_ProjectRegion2_2013 <- glmmTMB(counts ~ Year + ProjectRegion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)


nbinom_glmm_model_ProjectRegion3_2013 <- glmmTMB(counts ~ Year * ProjectRegion+Depth+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_ProjectRegion4_2013 <- glmmTMB(counts ~ Year * ProjectRegion+Habitat+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_Year_2013 <- glmmTMB(counts ~ as.numeric(Year) + (1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_habitat_2013 <- glmmTMB(counts ~ Year * Habitat+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)


nbinom_glmm_model_habitat2_2013 <- glmmTMB(counts ~ Year + Habitat+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_habitat3_2013 <- glmmTMB(counts ~ Year + Habitat+Subregion+(1|SiteID/StationID),
                                 family = nbinom2, 
                                 data = Total.octo_no_2012)

nbinom_glmm_model_null_2013 <- glmmTMB(counts ~  + 1,
                           family = nbinom2, 
                            data = Total.octo)
