---
title: "Modeling Florida octocoral trends: Biomass"
author: "Ronen liberman"
date: "8-06-2024"
output: 
  html_document:
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r initiate-environment, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library("knitr")
library("gridExtra")
library("ggpubr")
library("data.table")
library(glmmTMB)
library(bbmle)
library(performance)
library(ggplot2)
library(plotrix)
library(gridExtra)
library(car)
library(emmeans)
library(DHARMa)
library(RRPP)
library(MuMIn)
library(MASS)
library(lattice)
```

# Data preparation for modelling 
```{r}
#Use the density and heights by station data that was prepared using the "Creating_binding_all_datasets.R" script 
dat <- read.csv("~/Postdocing_NSU/Octocoral forests/Manuscript/scripts/processed_data/all_density_and_heights_by_stationID.csv")
head(dat)

# Creating a list of sites that should be omitted from the datasets 
omit_sites <- c("Martin County 1" ,"Martin County 2","Davis Rock","Palm Beach 1","Red Dun Reef","Palmata Patch","Loggerhead Patch", "Texas Rock","Broward County A")
species_biomass <- c( "Antillogorgia_americana", "Eunicea_flexuosa" ,"Gorgonia_ventalina")
# Assign data groups according to the frequncy found above. 


colnames(dat)
unique(dat$SPP_Code)
#glimpse(biomass_data)

# Step 1: Remove duplicates
cleaned_biomass_data <- dat %>%
  dplyr::select(-Date.x) %>%
  filter(sciName %in% species_biomass)  %>% 
  filter(!(Site_name %in% omit_sites)) %>% 
  distinct(Year, Subregion,SiteID, Site_name, StationID, SPP_Code, .keep_all = TRUE) %>% # make sure no duplicates in the data
  mutate(Depth_group = case_when(
    Depth < 21 ~ "Shallow",
    Depth >= 21 & Depth < 28 ~ "Mid-shallow", Depth >= 28 & Depth < 45  ~ "Mid-deep",
    Depth >= 45 ~ "Deep"
  ))

colnames(cleaned_biomass_data)

unique(cleaned_biomass_data$SPP_Code)

cleaned_biomass_data$Reg.Hab <- paste(cleaned_biomass_data$ProjectRegion, "_", cleaned_biomass_data$Habitat)
cleaned_biomass_data$Reg.Hab <- as.factor(cleaned_biomass_data$Reg.Hab)
cleaned_biomass_data$Habitat <- as.factor(cleaned_biomass_data$Habitat)
cleaned_biomass_data$Subregion <- as.factor(cleaned_biomass_data$Subregion)
cleaned_biomass_data$ProjectRegion <- as.factor(cleaned_biomass_data$ProjectRegion)
cleaned_biomass_data$Depth_group <- as.factor(cleaned_biomass_data$Depth_group ) 

# Step 2: Summarize total biomass per station
total_biomass_per_transect <- cleaned_biomass_data %>%
  filter(Year != 2012) %>%  # filter out the year 2012 
  group_by(Year, Subregion, Site_name, StationID) %>%
  summarize(
    pooled.taxa_biomass = sum(biomass, na.rm = TRUE),
    n = n()
  ) %>%
  ungroup()

# View the summarized biomass per station
print(total_biomass_per_transect)

# Extract unique station metadata
station_metadata <- cleaned_biomass_data %>%
  dplyr::select("SiteID","StationID", "ProjectRegion","Habitat","Reg.Hab", "Depth","Depth_group", "latDD", "lonDD", "latDeg", "latMin", "lonDeg", "lonMin") %>%
  distinct()

# Merge summarized biomass data with station metadata
all_biomass <- total_biomass_per_transect %>%
  left_join(station_metadata, by = c( "StationID")) %>% 
  rename( "biomass" = "pooled.taxa_biomass")

glimpse(all_biomass)

#Similar to the density analysis, here I remove sites that have focused specifically on other invertebrates, or that unnaturally had no octocorals. 
#omit_sites2 <- c("Red Dun Reef","Palmata Patch","Loggerhead Patch", "Texas Rock","Broward County A","Davis Rock")

# look at the frequency of biomass 

hist(all_biomass$biomass, breaks=50, xlab = "biomass", main = NULL, freq = TRUE)

#export
write.csv(file= "~/Postdocing_NSU/Octocoral forests/Manuscript/scripts/processed_data/processed_biomass_data.csv", all_biomass)
```
# Model selection - Check distribution

**Nbinom1 is preffrred** 
```{r echo=FALSE}
# if you start from here, use processed biomass_data.csv
#all_biomass <- read.csv("~/processed_biomass_data.csv")

poisson_glmm_full_model_2013 = glmmTMB(biomass ~ Year * Subregion +Depth+Habitat+(1|SiteID/StationID),
                                       family = poisson, 
                                       data = all_biomass)
poisson_glmm_full_model_2013 <- update(poisson_glmm_full_model_2013, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom1_glmm_full_model_2013 = glmmTMB(biomass ~ Year * Subregion +Depth+Habitat+(1|SiteID/StationID),
                                       family = nbinom1,
                                       data = all_biomass)
nbinom1_glmm_full_model_2013 <- update(nbinom1_glmm_full_model_2013, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_full_model_2013 = glmmTMB(biomass ~ Year * Subregion +Depth+Habitat+(1|SiteID/StationID),
                                      family = nbinom2, 
                                      data = all_biomass)
nbinom_glmm_full_model_2013 <- update(nbinom_glmm_full_model_2013, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

AIC(poisson_glmm_full_model_2013, nbinom1_glmm_full_model_2013, nbinom_glmm_full_model_2013)

```

Model selection - All possible models
```{r}


nbinom1_glmm_full_model_2013 = glmmTMB(biomass ~ Year * Subregion +Depth+Habitat+(1|SiteID/StationID),
                                       family = nbinom1,
                                       data = all_biomass1)
nbinom1_glmm_full_model_2013 <- update(nbinom1_glmm_full_model_2013, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depthgroup_interaction_2013 <- glmmTMB(biomass ~ Year * Depth+ (1|SiteID/StationID),
                                                         family = nbinom1, 
                                                         data = all_biomass)
#nbinom_glmm_model_depthgroup_interaction <- #update(nbinom_glmm_model_depthgroup_interaction, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depthgroup_2013 <- glmmTMB(biomass ~ Year * Depth_group + (1|SiteID/StationID),
                                             family = nbinom1, 
                                             data = all_biomass)
#nbinom_glmm_model_depthgroup <- update(nbinom_glmm_model_depthgroup, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_depth_2013 <- glmmTMB(biomass ~ Year + Depth + (1|SiteID/StationID),
                                        family = nbinom1, 
                                        data = all_biomass)

nbinom_glmm_model_Subregion_2013 <- glmmTMB(biomass ~ Year * Subregion+(1|SiteID/StationID),
                                            family = nbinom1, 
                                            data = all_biomass)

#nbinom_glmm_model_Subregion_2013 <- update(nbinom_glmm_model_Subregion_2013, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_Subregion2_2013 <- glmmTMB(biomass ~ Year + Subregion+(1|SiteID/StationID),
                                             family = nbinom1, 
                                             data = all_biomass)
#  nbinom_glmm_model_Subregion2 <- update(nbinom_glmm_model_Subregion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))                    

nbinom_glmm_model_Subregion3_2013 <- glmmTMB(biomass ~ Year * Subregion+Depth + (1|SiteID/StationID),
                                             family = nbinom1, 
                                             data = all_biomass)   
nbinom_glmm_model_Subregion3_2013 <- update(nbinom_glmm_model_Subregion3_2013, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))


nbionom_glmm_model_ProjectRegion_2013 <- glmmTMB(biomass ~ Year * ProjectRegion+(1|SiteID/StationID),
                                                 family = nbinom1, 
                                                 data = all_biomass)
#nbionom_glmm_model_ProjectRegion <- update(nbionom_glmm_model_ProjectRegion, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_ProjectRegion2_2013 <- glmmTMB(biomass ~ Year + ProjectRegion+(1|SiteID/StationID),
                                                 family = nbinom1, 
                                                 data = all_biomass)
#nbinom_glmm_model_ProjectRegion2 <- update(nbinom_glmm_model_ProjectRegion2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

nbinom_glmm_model_ProjectRegion3_2013 <- glmmTMB(biomass ~ Year * ProjectRegion+Depth+(1|SiteID/StationID),
                                                 family = nbinom1, 
                                                 data = all_biomass)

nbinom_glmm_model_ProjectRegion4_2013 <- glmmTMB(biomass ~ Year * ProjectRegion+Habitat+(1|SiteID/StationID),
                                                 family = nbinom1, 
                                                 data = all_biomass)

nbinom_glmm_model_Year_2013 <- glmmTMB(biomass ~ as.numeric(Year) + (1|SiteID/StationID),
                                       family = nbinom1, 
                                       data = all_biomass)
#nbinom_glmm_model_Year <- update(nbinom_glmm_model_Year, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))


nbinom_glmm_model_habitat_2013 <- glmmTMB(biomass ~ Year * Habitat+(1|SiteID/StationID),
                                          family = nbinom1, 
                                          data = all_biomass)


nbinom_glmm_model_habitat2_2013 <- glmmTMB(biomass ~ Year + Habitat+(1|SiteID/StationID),
                                           family = nbinom1, 
                                           data = all_biomass)

nbinom_glmm_model_habitat3_2013 <- glmmTMB(biomass ~ Year + Habitat+Subregion+(1|SiteID/StationID),
                                           family = nbinom1, 
                                           data = all_biomass)


nbinom_glmm_model_reg.hab_2013 <- glmmTMB(biomass ~ Year * Reg.Hab+(1|SiteID/StationID),
                                          family = nbinom1, 
                                          data = all_biomass)

nbinom_glmm_model_null_2013 <- glmmTMB(biomass ~  + 1,
                                       family = nbinom1, 
                                       data = all_biomass)

AICtab(nbinom1_glmm_full_model_2013,
       nbinom_glmm_model_depthgroup_interaction_2013,nbinom_glmm_model_depthgroup_2013,
       nbinom_glmm_model_depth_2013,
       nbinom_glmm_model_Subregion_2013,nbinom_glmm_model_Subregion2_2013, nbinom_glmm_model_Subregion3_2013,
       nbionom_glmm_model_ProjectRegion_2013,nbinom_glmm_model_ProjectRegion2_2013,
       nbinom_glmm_model_ProjectRegion3_2013,nbinom_glmm_model_ProjectRegion4_2013,
       nbinom_glmm_model_Year_2013,
       nbinom_glmm_model_habitat_2013,nbinom_glmm_model_habitat2_2013,nbinom_glmm_model_habitat3_2013,
       nbinom_glmm_model_null_2013, nbinom_glmm_model_reg.hab_2013)

BICtab(nbinom1_glmm_full_model_2013,
    nbinom_glmm_model_depthgroup_interaction_2013,nbinom_glmm_model_depthgroup_2013,
    nbinom_glmm_model_depth_2013,
    nbinom_glmm_model_Subregion_2013,nbinom_glmm_model_Subregion2_2013, nbinom_glmm_model_Subregion3_2013,
    nbionom_glmm_model_ProjectRegion_2013,nbinom_glmm_model_ProjectRegion2_2013,
    nbinom_glmm_model_ProjectRegion3_2013,nbinom_glmm_model_ProjectRegion4_2013,
    nbinom_glmm_model_Year_2013,
    nbinom_glmm_model_habitat_2013,nbinom_glmm_model_habitat2_2013,nbinom_glmm_model_habitat3_2013,
    nbinom_glmm_model_null_2013, nbinom_glmm_model_reg.hab_2013)
```

BEST MODEL = Year*Subregion+Depth , According to AIC and BIC. Notice, The full model is not far behind.
However, simplicity is also important and I will procedd with model_Subregion3.  

```{r echo=FALSE}
best_mod <- nbinom_glmm_model_Subregion3_2013                

summary(best_mod) # Depth is very significant here 

r2(nbinom_glmm_model_Subregion3_2013) #fit is very good with the random effects, and pretty high for the fixed factors too 

```

# Diagnosis of warnings
I recieved similar warnings on almost all models. Trying to assess their source is challenging. The most probable explanation I can find is that maybe there are some paparmeters that do not add to the results of the model ("suggest unnecessary components"). However this is strange as this model is better that simialr one without depth, or with only site as a random factor. 

The best model is still ok, no major errors were identifyied. 
```{r}
diagnose(best_mod) 
```
# Model Validation - **Visualizing residuals**

This model has issues with uniformity (ks test). Outliars also problematic, but not after testing using boostrapping. So really the Uniformity is an issue. 
```{r echo=FALSE}

#  Plotting residuals vs. fitted shows more or less the same result.
# Check for residual pattern within groups and difference between groups     
  xyplot(residuals(best_mod) ~ fitted(nbinom1_glmm_full_model_2013),
         panel=function(x, y){
           panel.xyplot(x, y)
           panel.loess(x, y, span = 3)
           panel.lmline(x, y, lty = 2)  # Least squares broken line
         }
  )

```


```{r}
simulationOutput <- simulateResiduals(best_mod, n = 1000, plot = TRUE,quantreg = T)
  testDispersion(simulationOutput) #dispersion fine, but just margnially 
  testOutliers(simulationOutput, type = "bootstrap")
  testResiduals(simulationOutput)
  plotResiduals(simulationOutput, all_biomass$SiteID, xlab = "Residuals:Site", main=NULL)
```

# Data visualization 
Looking at the data per site. 

Although I didnt plot Year here, I can not find major outliars in the raw data. maybe except for two sites. I check them as well, the data is ok. 

```{r}
to_keep <- c("ProjectRegion", "Habitat","Reg.Hab", "Depth","Depth_group") 


Biomass_per_site <- all_biomass %>%
  group_by(Year, Subregion, SiteID) %>%
  summarize(
    mean_biomass = mean(biomass, na.rm = TRUE),
    sd_biomass = sd(biomass, na.rm = TRUE),
    n = n(),
    across(all_of(to_keep), ~first(.)),
    .groups = "drop"
  )

head(Biomass_per_site)


p1 <- ggplot(Biomass_per_site, aes(x=SiteID,y = mean_biomass)) +
  geom_boxplot() +
  facet_wrap(~ Subregion, scales = "free_x") +
  labs(title = "Biomass across Different Sites", x = "Station ID", y = "Biomass") +
  theme_minimal()
p1

```

# Modeling with biomass averaged per site 
This might have reduced the uniformity by a little bit, but didnt resolve it completly. Still unsure this is what I need. 
 
```{r}
Biomass_per_site_filter <- Biomass_per_site %>% 
  filter(!(n <3)
)

glimpse(Biomass_per_site)
## 
mod_Subregion3_2013 <- glmmTMB(mean_biomass ~ Year * Subregion+Depth + (1|SiteID), 
                                             family = nbinom1, 
                                             data = Biomass_per_site)   

simulationOutput2 <- simulateResiduals(mod_Subregion3_2013, n = 1000, plot = TRUE,quantreg = T)
testDispersion(simulationOutput2) #dispersion fine, but just margnially 
testOutliers(simulationOutput2, type = "bootstrap")
testResiduals(simulationOutput2)

plotResiduals(simulationOutput2, Biomass_per_site$SiteID, xlab = "Residuals:Site", main=NULL)

```

