---
title: "Modelling_cover_data"
author: "Ronen liberman"
date: "2024-06-13"
output: 
  html_document:
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install_libraries, eval=FALSE}
# Install necessary libraries
install.packages(c("rmdformats", "prettydoc", "hrbrthemes", "tint", "tufte"))
```

```{r initiate-environment, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library("knitr")
library("gridExtra")
library("ggpubr")
library("data.table")
library(lme4)
library(glmmTMB)
library(bbmle)
library(performance)
library(ggplot2)
library(plotrix)
library(gridExtra)
library(car)
library(emmeans)
library(DHARMa)
library(RRPP)
library(MuMIn)
library(MASS)
library(lattice)
library(rmdformats)
library(prettydoc)
library(hrbrthemes)
library(tint)
library(tufte)
```


Data prep. 
I am using the "Cover_All.csv" dataset. The sites are similar to the list of sites that were used to model density and biomass data. 
A code chunk of data preparation should be added here later 

```{r}
Cover_dat <- read.csv("~/Postdocing_NSU/Octocoral forests/Manuscript/scripts/Cover_All.csv")
# Convert Depth to numeric
Cover_dat$Depth <- as.numeric(as.character(Cover_dat$Depth))
Cover_dat <- Cover_dat[,-1] #remove extra column

# Clean the Habitat column by trimming leading and trailing spaces
library(stringr)
Cover_dat <- Cover_dat %>%
  mutate(Habitat = str_trim(Habitat))
unique(Cover_dat$Habitat)

# Convert SiteID and StationID to factors 
Cover_Depth <- Cover_dat %>%
  mutate(SiteID = as.factor(SiteID),
         StationID = as.factor(StationID))%>% 
  mutate(Depth_group = case_when(
    Depth < 20 ~ "Shallow",
    Depth >= 20 & Depth < 45 ~ "Mid", 
    Depth >= 45 ~ "Deep"
  ))

#Add depth groups#
Cover_Depth <- Cover_dat %>% 
  mutate(Depth_group = case_when(
    Depth < 20 ~ "Shallow",
    Depth >= 20 & Depth < 45 ~ "Mid", 
    Depth >= 45 ~ "Deep"
  ))


glimpse(Cover_Depth)
unique(Cover_Depth$Subregion)
unique(Cover_Depth$SiteName)
unique(Cover_Depth$Habitat)
```
Modeling 
Here I follow the same model structure that has been developed for the density and biomass data. The difference here is that I run models using 'family = binomial', and the model takes into account the number of points per image as weights. 

```{r message=TRUE, warning=FALSE}
cover_full_model = glmmTMB(OCTO ~ factor(Year) * Subregion+Depth+Habitat+(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)


cover_depth <- glmmTMB(OCTO ~ factor(Year) * Depth+ (1|SiteID/StationID),
                          control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),
                       weights = points,family = binomial,data = Cover_Depth)

cover_depth2 <- glmmTMB(OCTO ~ factor(Year) * Depth_group + (1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_Subregion <- glmmTMB(OCTO ~ factor(Year) * Subregion+(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_Subregion2 <- glmmTMB(OCTO ~ factor(Year) + Subregion+Depth+(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_Subregion3 <- glmmTMB(OCTO ~ factor(Year) * Subregion+Depth_group+(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_Subregion4 <- glmmTMB(OCTO ~ factor(Year) + Subregion+Habitat+(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_ProjectRegion <- glmmTMB(OCTO ~ factor(Year) * ProjectRegion +(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_ProjectRegion2 <-  glmmTMB(OCTO ~ factor(Year) + ProjectRegion +Depth+(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)


cover_ProjectRegion3 <-  glmmTMB(OCTO ~ factor(Year) * ProjectRegion+Habitat+(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_Year <- glmmTMB(OCTO ~ factor(Year) * Depth+(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)


cover_habitat <- glmmTMB(OCTO ~ factor(Year) * Habitat+(1|SiteID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_habitat2 <- glmmTMB(OCTO ~ Year + Habitat+(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_habitat3 <- glmmTMB(OCTO ~ factor(Year) * Habitat+Depth+(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_habitat4 <- glmmTMB(OCTO ~ factor(Year) + Habitat+Depth_group +(1|SiteID/StationID), control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)

cover_null <- glmmTMB(OCTO ~  + 1, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)),weights = points,
                                                    family = binomial, 
                                                    data = Cover_Depth)
```

#Model selection
```{r echo=FALSE}
AICtab(cover_full_model,
       cover_depth,cover_depth2,
       cover_Subregion,cover_Subregion2,cover_Subregion3,cover_Subregion4,
       cover_ProjectRegion,cover_ProjectRegion2,cover_ProjectRegion3,
       cover_Year,
       cover_habitat,cover_habitat2,cover_habitat3,cover_habitat4,
       cover_null)

#BIC(cover_full_model,
 #      cover_depth,cover_depth2,
#       cover_Subregion,cover_Subregion2,cover_Subregion3,cover_Subregion4,
#       cover_ProjectRegion,cover_ProjectRegion2,cover_ProjectRegion3,
#       cover_Year,
#       cover_habitat,cover_habitat2,cover_habitat3,cover_habitat4,
#       cover_null)
```
The best model , with the lowest AIC was found to be cover_habitat3, this model has Year*Habitat + Depth as explanatory fixed factors and Station nested in Site as Random effect 
Model Summary 
```{r echo=FALSE}
summary(cover_habitat3)
```

Validation of the best model
Plotting residuals vs. fitted shows there ia high variation in the High numbers, similar to density data more or less

```{r echo=FALSE}
# Check for residual pattern within groups and difference between groups     
xyplot(residuals(cover_habitat3) ~ fitted(cover_full_model),
       panel=function(x, y){
         panel.xyplot(x, y)
         panel.loess(x, y, span = 3)
         panel.lmline(x, y, lty = 2)  # Least squares broken line
       }
)
```

Validation of the best model - cover_habitat3 using the DHarma package
```{r message=FALSE, warning=FALSE}
cover_simulationOutput <- simulateResiduals(cover_habitat3 , n = 2000, plot = TRUE,quantreg = T) #  Dispersion is OK, QQ plot is fine, but it seems this model has issues with ouliars and uniformity 


#Check individually 
testDispersion(cover_simulationOutput) #dispersion fine
testOutliers(cover_simulationOutput, type = "bootstrap") #OUTLIAR test significant: 0.02! This can be a problem!

#To see if we this model has outliars in specific groups I can look at the specific model explanatory variables 
# Check per Habitat 
cover_resids1 <- recalculateResiduals(cover_simulationOutput, group = Cover_All$Habitat)
testResiduals(cover_resids1)

# here we see that that if we look at the outlines per the habitat  group the result is no significant in any parameter, thus our model is validated for this group.  

# Check per Year
cover_resids2 <- recalculateResiduals(cover_simulationOutput, group = Cover_All$Year)
testResiduals(cover_resids2)
 # here we have a problem with the uniformity, which make sense. But dispersion and outliars are fine. 

# Check per Site
cover_resids3 <- recalculateResiduals(cover_simulationOutput, group = Cover_All$Site)
testResiduals(cover_resids3)

#Site reveals problem in KS and outliars 

# Check per Site
cover_resids4 <- recalculateResiduals(cover_simulationOutput, group = Cover_All$Depth)
testResiduals(cover_resids4)

 # here we see that that if we look at the outlines per the Depth group the result is no significant in any parameter, thus our model is validated for this group.  

```
