# Octocoral Juvenile 

This part of the script was prepared for analyzing the proportion of octocoral Juveniles on Florida's Coral reefs. The first chunk is for preparing data from raw (not running now). This code creats a data for to compare our 5 target species only using their repective densities and heights.Afterwards the code below cleans the data a little bit, then uses GLMMs with binomial distribution to test for best fit model using model selection. We examine the most significant parameters that influence state of octocoral Juveniles on in FCR during 2013-2023.

```{r Raw data prep., eval=FALSE, include=FALSE}
#Lets get data for density of 5 species , because so far I did it only for total octocoral density 


dat <- read.csv("~/Postdocing_NSU/Octocoral forests/Manuscript/scripts/processed_data/all_density_and_heights_by_stationID.csv")#view(Total.octo_no_2012)

density_DT<-  read.csv("~/Postdocing_NSU/Octocoral forests/Manuscript/scripts/processed_data/DTCREMP_OCTO_Summaries_2011-2023_Density.csv" ) 

density_SE <-  read.csv("~/Postdocing_NSU/Octocoral forests/Monitoring data/SECREMP_CSV_Files/SECREMP_OCTO_Summaries_2012-2023_Density.csv" ) 
density_CREMP <- read.csv("~/Postdocing_NSU/Octocoral forests/Monitoring data/CREMP_CSV_Files/CREMP_OCTO_Summaries_2011-2023_Density.csv" ) 


# Combine the datasets
combined_dens <- bind_rows(density_DT, density_SE, density_CREMP)
head(combined_dens, 20)

colnames(combined_dens)
#step 1: make long 
density_long_dat <-  combined_dens %>% 
  dplyr::select(-c(Eunicea_calyculata,Eunicea._calyculata )) %>% 
  pivot_longer(cols=c('Gorgonia_ventalina', 'Antillogorgia_americana', 'Eunicea_flexuosa','Antillogorgia_bipinnata', 'Pseudoplexaura_porosa' ,'Total_Octocorals'),
               names_to='sciName',
               values_to='density') %>% 
  filter(Year != 2011) 
# Creating a list of sites that should be omitted from the datasets 
omit_sites <- c("Martin County 1" ,"Martin County 2","Davis Rock","Palm Beach 1","Red Dun Reef","Palmata Patch","Loggerhead Patch", "Texas Rock","Broward County A")

species <- c('Gorgonia_ventalina', 'Antillogorgia_americana', 'Eunicea_flexuosa','Antillogorgia_bipinnata', 'Pseudoplexaura_porosa')
#three_species <- c('Gorgonia_ventalina', 'Antillogorgia_americana', 'Eunicea_flexuosa')

target_dat <- density_long_dat %>%
  #dplyr::select(-c(Eunicea_calyculata, Eunicea._calyculata)) %>%  
  filter(Year != 2012) %>%
  filter(!(Site_name %in% omit_sites)) %>% 
  filter(sciName %in%  species) %>%  #omit total octocorals
  # filter(sciName %in%  species) %>% #if filtering for 3 species
  group_by(Year, Subregion, Habitat,Site_name,StationID) %>%
  summarize(target_density = sum(density, na.rm = TRUE),
            n = n()) %>% 
  ungroup()

head(target_dat) 
# this is the density of 5 traget species in each of the transects in each year 

##################
#This part below is taking raw height data to generate proportion of juveniles from the total community of the 5 target species for each transect in each year.

Height_SE <- read.table("~/Postdocing_NSU/Octocoral forests/Manuscript/raw_data/SE_OCTO_Raw_Data.txt", header = TRUE, sep = ",", quote = "\"", na.strings = "")
Height_DT<-  read.table("~/Postdocing_NSU/Octocoral forests/Manuscript/raw_data/DT_OCTO_Raw_Data.txt", header = TRUE, sep = ",", quote = "\"", na.strings = "")
Height_CREMP <-  read.table("~/Postdocing_NSU/Octocoral forests/Manuscript/raw_data/FLK_OCTO_Raw_Data.txt", header = TRUE, sep = ",", quote = "\"", na.strings = "")

colnames(Height_SE)[colnames(Height_SE) == "Site_name"] <- "SiteName"

# bind height data 
Height_all <- rbind(Height_SE, Height_CREMP, Height_DT) 

head(Height_all,10)

# Count the number of unique SPP_Code
unique_spp_count <- Height_CREMP %>% 
  distinct(SPP_Code) 
unique_spp_count
#remove data on E. calyculata
Height_all <- filter(Height_all, SPP_Code != "ECAL" & SPP_Code != "PBIP/PKAL")

distinct(Height_all, SPP_Code) 

height_dat <- Height_all %>%
  filter(Year > 2012) %>% 
  distinct() %>% 
  mutate(sciName = dplyr::recode(sciName,
                                 'Pseudopterogorgia americana' = 'Antillogorgia_americana',
                                 "Eunicea flexuosa" = "Eunicea_flexuosa", 
                                 "Gorgonia ventalina" = "Gorgonia_ventalina", 
                                 "Pseudoplexaura porosa" = "Pseudoplexaura porosa", 
                                 "Pseudopterogorgia bipinnata" = 'Antillogorgia_bipinnata', 
                                 "Not Slimy - P. americana" = 'Antillogorgia_americana')) 

head(height_dat) 

#### Add station metadata ########### 

transect_meta <- read.csv("~/Postdocing_NSU/Octocoral forests/Manuscript/raw_data/StationMetadata_corrected_habitatid.csv")
head(transect_meta)

meta_dat <- transect_meta %>% 
  rename(StationID = stationid, SiteID = siteid, SiteName= sitename) 
head(meta_dat)

# Join the combined dataset with meta_dat based on SiteID and StationID
colnames(height_dat)

height_with_metadata <- left_join(height_dat, meta_dat, by = c("SiteID", "StationID","SiteName"))

# Display the first few rows of the new dataset

colnames(height_with_metadata)
unique(height_with_metadata$habitatid) 
unique(height_with_metadata$Habitat)  

height.meta.df <- height_with_metadata %>% 
  mutate(ProjectRegion = ifelse(ProjectRegion == "NPSDT/FKNMS", "NPSDT", ProjectRegion)) %>%
  mutate(ProjectRegion = ifelse(Subregion == "DT", "NPSDT", ProjectRegion)) %>%
  filter(!(SiteName %in% omit_sites)) %>% 
  rename(Depth = offshoreDepth) %>% 
  mutate(SiteID = as.factor(SiteID),
         StationID = as.factor(StationID),
         Reg.Hab = as.factor(paste(ProjectRegion, "_", habitatid)),
         ProjectRegion = as.factor(ProjectRegion))

unique(height.meta.df$ProjectRegion)
unique(height.meta.df$Reg.Hab)
glimpse(height.meta.df)

# Generate 'mature' column
height.meta.df <- height.meta.df %>%
  mutate(mature = ifelse(Height.cm. <= 5, "Juvenile", "Adult"))
unique(height.meta.df$sciName)

# Calculate ratio Juvenile/Adults per Year, ProjectRegion
three_target_species <- c( "Antillogorgia_americana", "Eunicea_flexuosa" ,"Gorgonia_ventalina")
rm(species)
# step 3: generate a ratio of juveniles/ adults for every station . 
#then summerize it by projectRegion 
colnames(height.meta.df)
#rm(percent_mature_station)
unique(height.meta.df$SiteName)

percent_mature_station <- height.meta.df %>%
  #filter(sciName %in% species) %>% 
  mutate(mature = ifelse(Height.cm. <= 5, "Juvenile", "Adult")) %>%
  # group_by(Year, ProjectRegion,StationID) %>%
  group_by(Year, Subregion, habitatid,SiteName,StationID) %>%
  filter(sum(mature == "Juvenile") <= sum(mature == "Adult")) %>%
  ungroup() %>%
  group_by(Year, ProjectRegion,Subregion, habitatid,SiteName,StationID, mature) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = mature, values_from = count, values_fill = 0) %>%
  mutate(Juv_proportion = Juvenile /(Juvenile+Adult)) %>%  #calaculate proportion of juveniles put of juveniles+adults
  ungroup()
glimpse(percent_mature_station)
unique(percent_mature_station$habitatid)

#Merge 
# Convert StationID in percent_mature_station to integer
percent_mature_station <- percent_mature_station %>%
  mutate(StationID = as.integer(as.character(StationID)))
# Perform the left join based on the specified columns
merged_dataset <- left_join(percent_mature_station, target_dat, 
                            by = c("Year", "Subregion", "StationID", "SiteName" = "Site_name"))
#add more metadata 

transect_meta <- read.csv("~/Postdocing_NSU/Octocoral forests/Manuscript/raw_data/StationMetadata_corrected_habitatid.csv")
head(transect_meta)

colnames(transect_meta)
colnames(merged_dataset)

sites <- merged_dataset$SiteName # make a list of the sites I am working with, this is used for subsetting 

meta_dat <- transect_meta %>% 
  rename(StationID = stationid, SiteID = siteid, SiteName = sitename, Depth = offshoreDepth) %>% 
  mutate(ProjectRegion = ifelse(ProjectRegion == "NPSDT/FKNMS", "NPSDT", ProjectRegion)) %>% 
  filter(SiteName %in% sites) %>% 
  mutate(ProjectRegion = ifelse(SiteName == "Black Coral Rock", "NPSDT", ProjectRegion)) #Black Coral Rock is in the DRY tortugas~~ 

head(meta_dat)

traget_with_metadata <- left_join(merged_dataset, meta_dat, by = c("ProjectRegion","SiteName","habitatid", "StationID"))

unique(traget_with_metadata$habitatid)

traget_with_metadata$target_n <- traget_with_metadata$Adult+traget_with_metadata$Juvenile #Number of colonies in each transect
traget_with_metadata$Reg.Hab <- paste(traget_with_metadata$ProjectRegion, "_", traget_with_metadata$habitatid)

#export
#write.csv(file= "~/Postdocing_NSU/Octocoral forests/Manuscript/scripts/processed_data/processed_juvenile and density of target species1.csv", traget_with_metadata) 
```

### Data prep. 
I am using the "processed_juvenile and density of target species.csv" dataset. This was prepared in the previous code chunk. The sites are similar to the list of sites that were used to model density and Cover data. 

```{r Data prep., echo=FALSE}
#load density and juvenile proportion data for the 5 target species 


FL.juv <- read.csv("~/Postdocing_NSU/Octocoral forests/Manuscript/scripts/processed_data/processed_juvenile and density of target species1.csv")

colnames(FL.juv)
FL.juv <- FL.juv %>% 
  dplyr::select("Year","ProjectRegion" ,"Subregion","SiteName","SiteID", "StationID", "habitatid","Reg.Hab", "Depth","Juv_proportion","target_density", "target_n" )

# all categories for the modeling should be as.factor
FL.juv$Year <- as.factor(FL.juv$Year)
FL.juv$Reg.Hab <- as.factor(FL.juv$Reg.Hab)
FL.juv$habitatid <- as.factor(FL.juv$habitatid)
FL.juv$Subregion <- as.factor(FL.juv$Subregion)
FL.juv$ProjectRegion <- as.factor(FL.juv$ProjectRegion)
FL.juv$SiteName <- as.factor(FL.juv$SiteName) 
FL.juv$SiteID <- as.factor(FL.juv$SiteID) 
FL.juv$StationID <- as.factor(FL.juv$StationID) 

#chaeck the data 
unique(FL.juv$habitatid)
unique(FL.juv$Reg.Hab)

# Identify the rows with NA values in the merged dataset
rows_with_na_ex1 <- FL.juv[rowSums(is.na(FL.juv)) > 0, ]

# Print the rows with NA values in the merged dataset
print(rows_with_na_ex1) # no rows with NA

glimpse(FL.juv)
str(FL.juv)
#summary(FL.juv)

# Open a new plotting window with specified dimensions
dev.new(width = 10, height = 8)

# Create the histogram
hist(FL.juv$Juv_proportion)

ggplot(FL.juv, aes(x = Year, y = Juv_proportion, colour = habitatid))+
  geom_point(stat = "identity", size = 1)+
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(~Reg.Hab)+
  theme_bw()+
  theme(axis.text.y = element_text(size = 12, hjust = 1, colour = "black"),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 16), 
        axis.text.x = element_text(size = 12, colour = "black",angle = 45, vjust = 1, hjust = 0.8),
        plot.margin = margin(rep(15, 4)), 
        axis.line = element_line(size = c(0.5, 0.5, 0.5, 0.5)), 
        panel.grid.major.x = element_blank(), 
        strip.text = element_text(size = 15, face = "bold.italic"), 
        strip.background = element_rect(size=0.5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        strip.text.y = element_blank())

```
# Modeling octocoral juveniles
Here I follow the same model structure that has been developed for the density and cover datasets. The difference here is that I run models using 'family = binomial', and the model takes into account the number of octocoral colonies (sum of target species) per transect as weights.

### Model selection 
I am using 'FL.juv' data and 'binomial' family
```{r}
#Running models
juv_full_model <- glmmTMB(Juv_proportion ~ Year + Subregion+Depth+Reg.Hab + 
                              (1 | SiteID/StationID),
                               family = binomial, weights = target_n,
                               data = FL.juv)

juv_Year <- glmmTMB(Juv_proportion ~ as.numeric(Year) + (1|SiteID/StationID),
                                  family = binomial, weights = target_n,
                                data = FL.juv)


juv_depth.interacation  <- glmmTMB(Juv_proportion ~ Year * Depth + (1 | SiteID/StationID),
                                family = binomial, weights = target_n,
                                data = FL.juv)



juv_Subregion1 <- glmmTMB(Juv_proportion ~ Year * Subregion + (1 | SiteID/StationID),
                                    family = binomial, weights = target_n,
                                    data = FL.juv)

juv_Subregion2 <- glmmTMB(Juv_proportion ~ Year + Subregion + (1 | SiteID/StationID),
                                     family = binomial, weights = target_n,
                                     data = FL.juv)

juv_Subregion_depth <- glmmTMB(Juv_proportion ~ Year * Subregion + Depth + (1 | SiteID/StationID),
                                     family = binomial, weights = target_n,
                                     data = FL.juv)

juv_Subregion4 <- glmmTMB(Juv_proportion ~ Year * Subregion + Reg.Hab + (1 | SiteID/StationID),
                                     family = binomial, weights = target_n,
                                     data = FL.juv)

juv_ProjectRegion1 <- glmmTMB(Juv_proportion ~ Year * ProjectRegion + (1 | SiteID/StationID),
                                        family = binomial, weights = target_n,
                                        data = FL.juv)

juv_ProjectRegion2 <- glmmTMB(Juv_proportion ~ Year + ProjectRegion + 
                                   (1 | SiteID/StationID),
                                   family = binomial, weights = target_n,
                                         data = FL.juv)

juv_ProjectRegion_depth <- glmmTMB(Juv_proportion ~ Year * ProjectRegion + 
                                     Depth + (1 | SiteID/StationID),
                                    family = binomial, weights = target_n,
                                    data = FL.juv)

juv_ProjectRegion4 <- glmmTMB(Juv_proportion ~ Year *ProjectRegion + Reg.Hab+
                                   (1 |SiteID/StationID),
                                      family = binomial, weights = target_n,
                                      data = FL.juv)

juv_reg_hab <- glmmTMB(Juv_proportion ~ Year * Reg.Hab + 
                         (1 | SiteID/StationID),
                         family = binomial, weights = target_n,
                          data = FL.juv)


juv_reg_hab2 <- glmmTMB(Juv_proportion ~ Year * Reg.Hab + Depth + 
                          (1 | SiteID/StationID),
                           family = binomial, weights = target_n,
                           data = FL.juv)
binomial_model_reg_hab_v2 <- update(binomial_model_reg_hab_v2, control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))
 
juv_null <- glmmTMB(Juv_proportion ~ 1,
                               family = binomial, weights = target_n,
                               data = FL.juv)


```

Model comparison
```{r Model comparison, echo=FALSE}
### Model Comparison with AIC
AICtab(juv_full_model,
       juv_Year,
       juv_depth.interacation,
       juv_Subregion1,
       juv_Subregion2,
       juv_Subregion_depth,
       juv_Subregion4,
       juv_ProjectRegion1,
       juv_ProjectRegion2,
       juv_ProjectRegion_depth,
       juv_ProjectRegion4,
       juv_reg_hab,
       juv_reg_hab2,
       juv_null)


#for exporting AIC numbers 

# comparing the models , 
AIC(binomial_full_model, 
       binomial_model_depth,
       binomial_model_Subregion, binomial_model_Subregion2, binomial_model_Subregion3, binomial_model_Subregion4,
       binomial_model_ProjectRegion, binomial_model_ProjectRegion2, binomial_model_ProjectRegion3,
       binomial_model_Year,
       binomial_model_habitat, binomial_model_habitat2, binomial_model_habitat3,
       binomial_model_reg_hab, binomial_model_reg_hab_v2,
       binomial_model_null)

```

best model is binomial_model_reg_hab, but there are issues  

Best model summary
```{r}
summary(juv_reg_hab) 
summary(juv_reg_hab2) 
r2(juv_reg_hab) #0.145; 0.088 - most variation wxplained by fixed factors Year and Reg.Hab

r2(juv_reg_hab2) #0.145; 0.088 - most variation wxplained by fixed factors Year and Reg.Hab

```

### Model validation 
```{r Model validation , echo=FALSE}
best_mod <- juv_reg_hab

simulationOutput <- simulateResiduals(best_mod, n = 1000, plot = TRUE,quantreg = T)
# best Model looks good, no significant issues

testDispersion(simulationOutput) #dispersion fine

resids.site <- recalculateResiduals(simulationOutput, group = FL.juv$SiteName)
plot(resids.site) # heterogeneity good, uniformity good, no outliars

testResiduals(resids.site) #outliars very close to being significant, some evidence of outliers, try bootstrap as per Florian Hartig vignette advice

testOutliers(resids.site, type = "bootstrap") #suggests after bootstrapping that residuals p.value = 1, this is strange, but it also wasnt significant before. so still fine
testDispersion(resids.site) #dispersion fine

```

### Post-hoc
calculated estimated marginal means and outputs estimates on the log scale.
significant positive ratio estimate means juvenile prop. in year x is greater than in year y, significant negative means sig decline between year x and year y.
```{r Post hoc Juvenile, echo=FALSE}

emm.year.hab <- emmeans(best_mod, ~ Year|Reg.Hab) #compares years within habitat
pairs(emm.year.hab) 
contrast(emm.year.hab) #compares estimate for the year against the mean for that Reg.habitat
plot(emm.year.hab)
```
### Visualize proportion of juveniles / density
```{r Visualize, echo=FALSE}

# Calculate the average, standard deviation, and standard error
glimpse(FL.juv)
summary_stats <- FL.juv %>%
  group_by(Year, ProjectRegion) %>%
  summarize(mean_juv_prop = mean(Juv_proportion ),
            sd_prop = sd(Juv_proportion),
            se_prop = sd(Juv_proportion) / sqrt(n()),
            mean_target_density = mean(target_density),
            sd_target_density = sd(target_density),
            se_target_density = sd(target_density) / sqrt(n()))


head(summary_stats)

region_labels <- c(SECREMP = "SE Florida",
                   FKNMS = "Florida Keys",
                   NPSDT = "Dry Tortugas"
)

# Assuming 'summary_stats' is the summarized dataset
coeff_ratio <- 100  # Assuming 100 to scale the ratio for visualization

p_density_ratio <- ggplot(summary_stats, aes(x = Year, group = ProjectRegion)) +
  
  # Add line for mean target density
  geom_line(aes(y = mean_target_density, color = "Mean Target Density"),
            linewidth = 2, alpha = 0.8) +
  
  # Add ribbon for standard error of mean target density
  geom_ribbon(aes(ymin = mean_target_density - se_target_density,
                  ymax = mean_target_density + se_target_density,
                  ), 
              fill = "skyblue", alpha = 0.3) +
  
  # Add points and error bars for mean juvenile proportion (scaled)
  geom_point(aes(y = mean_juv_prop * coeff_ratio, color = "Juvenile/Adult Ratio"), 
             size = 3) +
  
  geom_errorbar(aes(ymin = (mean_juv_prop - se_prop) * coeff_ratio,
                    ymax = (mean_juv_prop + se_prop) * coeff_ratio, 
                    color = "Juvenile/Adult Ratio"),
                width = 0.2, alpha = 0.7) +
  
  scale_y_continuous(
    name = expression(bold("Mean density of target species (m"^2*")")),
    sec.axis = sec_axis(~ . / coeff_ratio, name = "Juvenile/Adult Ratio (%)")
  ) +
  
  scale_x_discrete(limits = levels(summary_stats$Year)) +
  
  scale_color_manual(values = c("Mean Target Density" = "dodgerblue3", 
                                "Juvenile/Adult Ratio" = "darkorange2"),name = "") +
  #scale_fill_manual(values = c("Mean Target Density" = "skyblue")) +
  
  facet_wrap(~ ProjectRegion,labeller = as_labeller(region_labels)) +
  
  theme_pubclean() +
  
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(margin = margin(t = 10), face = "bold"),
    axis.title.y = element_text(color = "black", face = "bold", size = 12, margin = margin(r = 10)),
    axis.title.y.right = element_text(color = "black", face = "bold", size = 12, margin = margin(l = 10)),
    strip.background = element_rect(fill = "white", color = NA),
    strip.text = element_text(face = "bold", size = 12),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "whitesmoke", color = NA),
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.margin = margin(20, 20, 20, 20)
  ) +
  
  labs(x = "Year", title = "Mean Target Species Density and proportions of their juveniles") +
  
  coord_cartesian(expand = FALSE)

# Print the plot
p_density_ratio
#export 
#ggsave(file ="~/Mean target species density and juvenile proportion.jpg", p_density_ratio, width =45 , height = 18,dpi=600,  units = "cm" )   

```

